---
title: "MasterVS_1"
author: "Kristian Suszczenia"
date: "2024-11-09"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Install the packages, and library. Library the other dependencies.
```{r}
library(virtualspecies)
library(ade4)
library(dismo)
library(rworldmap)
library(terra)
library(raster)
library(dplyr)
```


## 1) Load Input Layers

```{r}
#Load as SpatRaster
Elevation <- rast("C:/Users/shil5756/Documents/Directories/R_directories/Masters_VS_borneo/Elev_nibbled_jenk.tif")
MeanTemp <- rast("C:/Users/shil5756/Documents/Directories/R_directories/Masters_VS_borneo/BIO1_30s_now.tif")
AnnualPrec <- rast("C:/Users/shil5756/Documents/Directories/R_directories/Masters_VS_borneo/BIO12_30s_now.tif")

#Make our objects into a *string of layers*, and set more useful names
EnvLayers <- c(Elevation, MeanTemp, AnnualPrec)
names(EnvLayers) <- paste0(c("Elevation", "MeanTemp", "AnnualPrec"))
EnvLayers
```

We can see that they are the correct format (SpatRaster), sharing the same extent and resolution, and CRS.

## 2) Initial Suitability functions

For reference. We plan to create 6 species, 2 sets of 3. In each set there will be a Generalist, Wet Specialist, and Dry specialist. There will be a set for High and Low Elevation (so 6 in total).

# The parameters are as follows:

Elevation (range)

Highland = 479 to 3966
Lowland  = -36 to 478

Temperature (mean)

Highland mean = 23.07 (range; 8-28 sd; 1.573)
Lowland mean  = 26.73 (range; 22-29, sd; 0.741)

Precipitation (mean)

Highland dry = 2601.74 (range; 1630-3144, sd; 347.097) *HD*
Highland wet = 3687.75 (range; 3145-4252, sd; 282.767) *HW*
Highland Gen = 3341.10 (range; 1637-4252, sd; 590.927) *HG*

Lowland dry  = 2496.18 (range; 1572-2957, sd; 314.665) *LD*
Lowland wet  = 3420.59 (range; 2958-4310, sd; 331.150) *LW*
Lowland Gen  = 2955.42 (range; 1572-4310, sd; 563.852) *LG*

# Example: Highland DRY Specialist

prefers temperatures of 23.07* (dnorm), precipitation around 2601.74 l/yr (dnorm), and elevations above 479m from sea level (sigmoidal relationship is most appropriate). 

```{r}
#Set the env-suitability relationship functions, the collection of functions is the Initial suitability function (ISF)
ISF_HD <- formatFunctions(MeanTemp = c(fun = 'dnorm', mean = 23.07, sd = 1.573),
                          AnnualPrec = c(fun = 'dnorm', mean = 2601.74, sd = 347.097),
                          Elevation = c(fun = 'logisticFun', beta = 479, alpha = -70))

#These are then combined via a method (default is multiplicative) to create a scale of Habitat suitability scaling 0-1. The combination of functions is the Habitat suitability Index (HSI)
HSI_HD <- generateSpFromFun(raster.stack = EnvLayers[[c("MeanTemp", "AnnualPrec", "Elevation")]], 
                                      parameters = ISF_HD, 
                                      plot = TRUE)

#Plot to see how each function is operating in the HSI
plotResponse(HSI_HD)
```

Now we can convert this to a probability occurrence (which we will do linearly, so 0.6 means a 60% probability of occurrence). This is much better than the threshold method.

We should use a consistant way to set prevelence; such as 5% of the total land area pixels. This means each VS should end up with the same number of presences (or similar due to R setting the funciton to try and get as close to 5% as possible given the restrictions).

We also need to use a consistant alpha value for the logistic transformation, so that the shape of the probabiltiy of occurence is the same for each VS; meaning the only thing that changes is their preferences.

[DEPRECATED]

```{r}
Presence_HD <- convertToPA(HSI_HD, 
                   PA.method = "probability", #here you choose threshold or probability method
                   prob.method = "linear",
                   species.prevalence = 0.03, #defines the percentage of land area that will have presence points, thus the software will find the ideal linear relationship that satisfies that
                   plot = FALSE) 

plotSuitabilityToProba(Presence_HD)
plot(Presence_HD$pa.raster)
plot(Presence_HD$suitab.raster)
```


### Used VS:

#Highland DRY Specialist

```{r}
#Set the env-suitability relationship functions, the collection of functions is the Initial suitability function (ISF)
ISF_HD <- formatFunctions(MeanTemp = c(fun = 'dnorm', mean = 23.07, sd = 1.573),
                          AnnualPrec = c(fun = 'dnorm', mean = 2601.74, sd = 347.097),
                          Elevation = c(fun = 'logisticFun', beta = 479, alpha = -70))

#These are then combined via a method (default is multiplicative) to create a scale of Habitat suitability scaling 0-1. The combination of functions is the Habitat suitability Index (HSI)
HSI_HD <- generateSpFromFun(raster.stack = EnvLayers[[c("MeanTemp", "AnnualPrec", "Elevation")]], 
                                      parameters = ISF_HD, 
                                      plot = TRUE)

#Plot to see how each function is operating in the HSI
plotResponse(HSI_HD)

#Convert tor presence points
Presence_HD <- convertToPA(HSI_HD, 
                   PA.method = "probability", #here you choose threshold or probability method
                   prob.method = "linear", a = 1, b = 0, 
                   plot = FALSE) 

plotSuitabilityToProba(Presence_HD)
plot(Presence_HD$pa.raster)
```


#Highland WET Specialist
```{r}
#Set the env-suitability relationship functions, the collection of functions is the Initial suitability function (ISF)
ISF_HW <- formatFunctions(MeanTemp = c(fun = 'dnorm', mean = 23.07, sd = 1.573),
                          AnnualPrec = c(fun = 'dnorm', mean = 3687.75, sd = 282.767),
                          Elevation = c(fun = 'logisticFun', beta = 479, alpha = -70))

#These are then combined via a method (default is multiplicative) to create a scale of Habitat suitability scaling 0-1. The combination of functions is the Habitat suitability Index (HSI)
HSI_HW <- generateSpFromFun(raster.stack = EnvLayers[[c("MeanTemp", "AnnualPrec", "Elevation")]], 
                                      parameters = ISF_HW, 
                                      plot = TRUE)

#Plot to see how each function is operating in the HSI
plotResponse(HSI_HW)

Presence_HW <- convertToPA(HSI_HW, 
                   PA.method = "probability", 
                   prob.method = "linear", a = 1, b = 0, 
                   plot = FALSE) 

plotSuitabilityToProba(Presence_HW)
plot(Presence_HW$pa.raster)
```

### High Elevation Generalist
```{r}
#Set the env-suitability relationship functions, the collection of functions is the Initial suitability function (ISF)
ISF_HG <- formatFunctions(MeanTemp = c(fun = 'dnorm', mean = 23.07, sd = 1.573),
                          AnnualPrec = c(fun = 'dnorm', mean = 3341.1, sd = 590.927),
                          Elevation = c(fun = 'logisticFun', beta = 479, alpha = -70))

#These are then combined via a method (default is multiplicative) to create a scale of Habitat suitability scaling 0-1. The combination of functions is the Habitat suitability Index (HSI)
HSI_HG <- generateSpFromFun(raster.stack = EnvLayers[[c("MeanTemp", "AnnualPrec", "Elevation")]], 
                                      parameters = ISF_HG, 
                                      plot = TRUE)

#Plot to see how each function is operating in the HSI
plotResponse(HSI_HG)

#Convert to presence points
Presence_HG <- convertToPA(HSI_HG, 
                   PA.method = "probability", 
                   prob.method = "linear", a = 1, b = 0, 
                   plot = FALSE) 

plotSuitabilityToProba(Presence_HG)
plot(Presence_HG$pa.raster)
```

### Low Elevation DRY specialist
```{r}
#Set the env-suitability relationship functions, the collection of functions is the Initial suitability function (ISF)
ISF_LD <- formatFunctions(MeanTemp = c(fun = 'dnorm', mean = 26.73, sd = 0.741),
                          AnnualPrec = c(fun = 'dnorm', mean = 2496.18, sd = 314.665),
                          Elevation = c(fun = 'logisticFun', beta = 479, alpha = 70))

#These are then combined via a method (default is multiplicative) to create a scale of Habitat suitability scaling 0-1. The combination of functions is the Habitat suitability Index (HSI)
HSI_LD <- generateSpFromFun(raster.stack = EnvLayers[[c("MeanTemp", "AnnualPrec", "Elevation")]], 
                                      parameters = ISF_LD, 
                                      plot = TRUE)

#Plot to see how each function is operating in the HSI
plotResponse(HSI_LD)

#Convert to presence points
#Or logistic?
Presence_LD <- convertToPA(HSI_LD, 
                   PA.method = "probability", 
                   prob.method = "logistic", alpha = -0.08, 
                   species.prevalence = 0.1,
                   plot = FALSE) 

plotSuitabilityToProba(Presence_LD)
plot(Presence_LD$pa.raster)
```

### Low Elevation WET specialist
```{r}
#Set the env-suitability relationship functions, the collection of functions is the Initial suitability function (ISF)
ISF_LW <- formatFunctions(MeanTemp = c(fun = 'dnorm', mean = 26.73, sd = 0.741),
                          AnnualPrec = c(fun = 'dnorm', mean = 3420.59, sd = 331.15),
                          Elevation = c(fun = 'logisticFun', beta = 479, alpha = 70))

#These are then combined via a method (default is multiplicative) to create a scale of Habitat suitability scaling 0-1. The combination of functions is the Habitat suitability Index (HSI)
HSI_LW <- generateSpFromFun(raster.stack = EnvLayers[[c("MeanTemp", "AnnualPrec", "Elevation")]], 
                                      parameters = ISF_LW, 
                                      plot = TRUE)

#Plot to see how each function is operating in the HSI
plotResponse(HSI_LW)

#Convert to presence points
#or logistic
Presence_LW <- convertToPA(HSI_LW, 
                   PA.method = "probability", 
                   prob.method = "logistic", alpha = -0.08, 
                   species.prevalence = 0.1,
                   plot = FALSE) 

plotSuitabilityToProba(Presence_LW)
plot(Presence_LW$pa.raster)
```


## Low Elevation Generalist
```{r}
#Set the env-suitability relationship functions, the collection of functions is the Initial suitability function (ISF)
ISF_LG <- formatFunctions(MeanTemp = c(fun = 'dnorm', mean = 26.73, sd = 0.741),
                          AnnualPrec = c(fun = 'dnorm', mean = 2955.42, sd = 563.852),
                          Elevation = c(fun = 'logisticFun', beta = 479, alpha = 70))

#These are then combined via a method (default is multiplicative) to create a scale of Habitat suitability scaling 0-1. The combination of functions is the Habitat suitability Index (HSI)
HSI_LG <- generateSpFromFun(raster.stack = EnvLayers[[c("MeanTemp", "AnnualPrec", "Elevation")]], 
                                      parameters = ISF_LG, 
                                      plot = TRUE)

#Plot to see how each function is operating in the HSI
plotResponse(HSI_LG)

#Convert to presence points
#Or logsitic?
Presence_LG <- convertToPA(HSI_LG, 
                   PA.method = "probability", 
                   prob.method = "logistic", alpha = -0.08, 
                   species.prevalence = 0.1,
                   plot = FALSE) 

plotSuitabilityToProba(Presence_LG)
plot(Presence_LG$pa.raster)
```
################################################################################

### Sampling Presence/Absence

From the True Presence data, we can sample what will ultimately be the PA which can make up our habitat suitability model. 

# HD

```{r}
SampledPresence_HD <- sampleOccurrences(Presence_HD,
                            n = 5500, # The number of points to sample
                            type = "presence-absence",
                            sample.prevalence = 500/5500)

```

# HW
```{r}
SampledPresence_HW <- sampleOccurrences(Presence_HW,
                            n = 5500, # The number of points to sample
                            type = "presence-absence",
                            sample.prevalence = 500/5500)

```

# HG
```{r}
SampledPresence_HG <- sampleOccurrences(Presence_HG,
                            n = 5500, # The number of points to sample
                            type = "presence-absence",
                            sample.prevalence = 500/5500)

```

# LD
```{r}
SampledPresence_LD <- sampleOccurrences(Presence_LD,
                            n = 5500, # The number of points to sample
                            type = "presence-absence",
                            sample.prevalence = 500/5500)

```

# LW
```{r}
SampledPresence_LW <- sampleOccurrences(Presence_LW,
                            n = 5500, # The number of points to sample
                            type = "presence-absence",
                            sample.prevalence = 500/5500)

```

# LG
```{r}
SampledPresence_LG <- sampleOccurrences(Presence_LG,
                            n = 5500, # The number of points to sample
                            type = "presence-absence",
                            sample.prevalence = 500/5500)

SampledPresence_LG

```

###############################################################################

### Saving for Export to ARCGIS

This is just to showcase the process. The True Habitat suitability is transformed into a smaller population distribution liklihood (which puts them in the most suitable places, with comparable populations sizes), and these pixel probabilities are used to radnomly draw a Presence Absence True Population. 
```{r}
plot(Presence_LW$suitab.raster)
plot(Presence_LW$probability.of.occurrence)
plot(Presence_LW$pa.raster)
```
We will save the VS ('Presence_X') as this contains the habitat suitability raster and the true presence/absence. 
We will also save the suitability raster itself, separately, and the true presence absence, as TIFs.
Then we will save the sampled presence points as a csv.

```{r}

#HD
save(Presence_HD, file = "C:/Users/shil5756/Documents/Directories/R_directories/Masters_VS_borneo/Saved2/Saved_VS/VS_HD1")

writeRaster(Presence_HD$suitab.raster, 
            "C:/Users/shil5756/Documents/Directories/R_directories/Masters_VS_borneo/Saved2/Saved_realhabitatsuitability/Suitability_HD1.tif")
writeRaster(Presence_HD$pa.raster, 
            "C:/Users/shil5756/Documents/Directories/R_directories/Masters_VS_borneo/Saved2/Saved_realpresence/RealDist_HD1.tif")
write.csv(SampledPresence_HD$sample.points, 
          "C:/Users/shil5756/Documents/Directories/R_directories/Masters_VS_borneo/Saved2/Saved_samplepresence/P_HD1.csv")

#HW
save(Presence_HW, file = "C:/Users/shil5756/Documents/Directories/R_directories/Masters_VS_borneo/Saved2/Saved_VS/VS_HW1")

writeRaster(Presence_HW$suitab.raster, 
            "C:/Users/shil5756/Documents/Directories/R_directories/Masters_VS_borneo/Saved2/Saved_realhabitatsuitability/Suitability_HW1.tif")
writeRaster(Presence_HW$pa.raster, 
            "C:/Users/shil5756/Documents/Directories/R_directories/Masters_VS_borneo/Saved2/Saved_realpresence/RealDist_HW1.tif")
write.csv(SampledPresence_HW$sample.points, 
          "C:/Users/shil5756/Documents/Directories/R_directories/Masters_VS_borneo/Saved2/Saved_samplepresence/P_HW1.csv")

#HG

save(Presence_HG, file = "C:/Users/shil5756/Documents/Directories/R_directories/Masters_VS_borneo/Saved2/Saved_VS/VS_HG1")

writeRaster(Presence_HG$suitab.raster, 
            "C:/Users/shil5756/Documents/Directories/R_directories/Masters_VS_borneo/Saved2/Saved_realhabitatsuitability/Suitability_HG1.tif")
writeRaster(Presence_HG$pa.raster, 
            "C:/Users/shil5756/Documents/Directories/R_directories/Masters_VS_borneo/Saved2/Saved_realpresence/RealDist_HG1.tif")
write.csv(SampledPresence_HG$sample.points, 
          "C:/Users/shil5756/Documents/Directories/R_directories/Masters_VS_borneo/Saved2/Saved_samplepresence/P_HG1.csv")

#LD
save(Presence_LD, file = "C:/Users/shil5756/Documents/Directories/R_directories/Masters_VS_borneo/Saved2/Saved_VS/VS_LD1")

writeRaster(Presence_LD$suitab.raster, 
            "C:/Users/shil5756/Documents/Directories/R_directories/Masters_VS_borneo/Saved2/Saved_realhabitatsuitability/Suitability_LD1.tif")
writeRaster(Presence_LD$pa.raster, 
            "C:/Users/shil5756/Documents/Directories/R_directories/Masters_VS_borneo/Saved2/Saved_realpresence/RealDist_LD1.tif")
write.csv(SampledPresence_LD$sample.points, 
          "C:/Users/shil5756/Documents/Directories/R_directories/Masters_VS_borneo/Saved2/Saved_samplepresence/P_LD1.csv")

#LW
save(Presence_LW, file = "C:/Users/shil5756/Documents/Directories/R_directories/Masters_VS_borneo/Saved2/Saved_VS/VS_LW1")

writeRaster(Presence_LW$suitab.raster, 
            "C:/Users/shil5756/Documents/Directories/R_directories/Masters_VS_borneo/Saved2/Saved_realhabitatsuitability/Suitability_LW1.tif")
writeRaster(Presence_LW$pa.raster, 
            "C:/Users/shil5756/Documents/Directories/R_directories/Masters_VS_borneo/Saved2/Saved_realpresence/RealDist_LW1.tif")
write.csv(SampledPresence_LW$sample.points, 
          "C:/Users/shil5756/Documents/Directories/R_directories/Masters_VS_borneo/Saved2/Saved_samplepresence/P_LW1.csv")

#LG
save(Presence_LG, file = "C:/Users/shil5756/Documents/Directories/R_directories/Masters_VS_borneo/Saved2/Saved_VS/VS_LG1")

writeRaster(Presence_LG$suitab.raster, 
            "C:/Users/shil5756/Documents/Directories/R_directories/Masters_VS_borneo/Saved2/Saved_realhabitatsuitability/Suitability_LG1.tif")
writeRaster(Presence_LG$pa.raster, 
            "C:/Users/shil5756/Documents/Directories/R_directories/Masters_VS_borneo/Saved2/Saved_realpresence/RealDist_LG1.tif")
write.csv(SampledPresence_LG$sample.points, 
          "C:/Users/shil5756/Documents/Directories/R_directories/Masters_VS_borneo/Saved2/Saved_samplepresence/P_LG1.csv")
```

### Taking notes

We can also count the population size, for documentation
```{r}
#HD
HD_values <- as.data.frame(na.omit(values(Presence_HD$pa.raster)))
HD_values <- HD_values %>% rename_at('VSP suitability', ~'PA')
HD_values

nrow(HD_values) # Number of pixels overall = 858,655
sum(HD_values$PA == 1) # HD presence pixels = 25,988

#HG
HG_values <- as.data.frame(na.omit(values(Presence_HG$pa.raster)))
HG_values <- HG_values %>% rename_at('VSP suitability', ~'PA')
HG_values

nrow(HG_values)
sum(HG_values$PA == 1) # HG presence pixels = 79,681

#HW
HW_values <- as.data.frame(na.omit(values(Presence_HW$pa.raster)))
HW_values <- HW_values %>% rename_at('VSP suitability', ~'PA')
HW_values

nrow(HW_values)
sum(HW_values$PA == 1) # HW presence pixels = 55,894

#LD
LD_values <- as.data.frame(na.omit(values(Presence_LD$pa.raster)))
LD_values <- LD_values %>% rename_at('lyr.1', ~'PA')
LD_values

nrow(LD_values)
sum(LD_values$PA == 1) # LD presence pixels = 86,638

#LG
LG_values <- as.data.frame(na.omit(values(Presence_LG$pa.raster)))
LG_values <- LG_values %>% rename_at('lyr.1', ~'PA')
LG_values

nrow(LG_values)
sum(LG_values$PA == 1) # LG presence pixels = 85,777

#LW
LW_values <- as.data.frame(na.omit(values(Presence_LW$pa.raster)))
LW_values <- LW_values %>% rename_at('lyr.1', ~'PA')
LW_values

nrow(LW_values)
sum(LW_values$PA == 1) # LG presence pixels = 85,818

```


###############################################################################









