---
title: "Change_topog_corre"
author: "Kristian Suszczenia"
date: "2024-10-29"
output: html_document
---

# Preliminary Analysis

This rmd file entails the preliminary finding that guided virtual species creation; that precipition alteration due to climate change increases with elevation, and so does temperature to a lesser extent. This was used to guide the specialism and generalism of VS. 

The main goal is to load in the temperature and precipitation change rasters provided (which I made by subtracting the future from the current) and testing if this correlates with any topology in the landscape. 

As in the rest of this repo, #explanations will only be given on first instance of code use

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Install and Library packages

```{r load pakcages}
#use install.packages if needed
library(terra)
library(ggplot2)
```

##1) Load input climate and topology Rasters

CTI raster
```{r load CTI}
CTI <- rast(here('data', 'PreliminaryData', 'CTI', 'CTI_5.tif'))
CTI
```
TPI raster
```{r Load TPI}
TPI <- rast(here('data', 'PreliminaryData', 'TPI', 'TPI_6.tif'))
TPI
```
Roughness raster
```{r Load roughness}
Ruff <- rast(here('data', 'PreliminaryData', 'Ruff', 'Ruff_5.tif'))
Ruff
```
Elevation raster
```{r load elevation}
#load file
Elev <- rast(here('data', 'PreliminaryData', 'Elevation', 'Elevation.asc'))

#view
Elev
plot(Elev)

#we may want to also have this in log data
logElev <- log1p(Elev) #use log +1 as there are zero values
plot(logElev)
```

Now the Temp change rasters
```{r load temp change rasters}
#load the temp change raster for ssp245
TempChange_ssp245_2040 <- rast(here('data', 'PreliminaryData', 'TempChange', 'TempChange245_2040norm.tif'))
#load the temp change raster for ssp370
TempChange_ssp370_2040 <- rast(here('data', 'PreliminaryData', 'TempChange', 'TempChange370_2040norm2.tif'))

#view
TempChange_ssp245_2040
TempChange_ssp370_2040
```
And Precipitation Change rasters
```{r load prec change rasters}
#load the prec change raster for ssp245
PrecChange_ssp245_2040 <- rast(here('data', 'PreliminaryData', 'PrecChange', 'PrecChange245_2040norm.tif'))
#load the prec change raster for ssp370
PrecChange_ssp370_2040 <- rast(here('data', 'PreliminaryData', 'PrecChange', 'PrecChange370_2040norm.tif'))

#view
PrecChange_ssp245_2040
PrecChange_ssp370_2040
```

## 2) Explore Temperature climate change correlations

Can use histograms to observe the distribution
```{r Histogram of Data}
#use par() to display side by side
par(mfrow = c(3, 3))
hist(PrecChange_ssp245_2040, #uses consistant number of breaks
     breaks = 50,
     main= "Precip Change ssp245")
hist(PrecChange_ssp370_2040, 
     breaks= 50,
     main= "Precip Change ssp370")
hist(TempChange_ssp245_2040, 
     breaks= 50,
     main= "Temp Change ssp245")
hist(TempChange_ssp370_2040, 
     breaks= 50,
     main= "Temp Change ssp370")
hist(CTI,
     breaks= 50,
     main= "CTI")
hist(TPI, 
     breaks= 50,
     main= "TPI")
hist(Ruff, 
     breaks= 50,
     main= "Roughness")
hist(logElev,
     breaks= 50,
     main= "Log Elevation")
hist(Elev,
     breaks= 50,
     main= "Elevation")
```
Or use plots, I will set the transparency very low in order to see the trend within alot of data

I have crated this function for multiple use;
```{r Load plotting function}
#this function will plot a correlation between desired variables, 
#as well as test a specified statistic
explore_cor <- function(x, y, x_name, y_name, cor.method){
  Lr <- lm(values(y) ~ values(x))
  
  plot(values(x), values(y),
     xlab= x_name,
     ylab= y_name,
     cex= 0.4,
     col = rgb(red = 0, green = 0, blue = 1, alpha = 0.005)
     ) +
  abline(Lr_1, col="black")
  
  cor.test(values(x), values(y), method= cor.method) 

}
```

#### a) Temp Change ssp245: Elevation
```{r Does Temperature scale with elevation?}
explore_cor(x= logElev, y= TempChange_ssp245_2040, 
            x_name= "log transformed Elevation", 
            y_name= "Temperature Change, Current - 2041 ssp245", 
            cor.method = 'spearman')

# rho = -0.2703917, p-value < 2.2e-16, alternative hypothesis: true rho is not equal to 0
```

#### b) Temp Change ssp245: CTI
```{r Does Temperature scale with Wetness Index?}
explore_cor(x= CTI, y= TempChange_ssp245_2040, 
            x_name= "Wetness Index (CTI)", 
            y_name= "Temperature Change, Current - 2041 ssp245", 
            cor.method = 'spearman')

# rho = -0.2703917, p-value < 2.2e-16, alternative hypothesis: true rho is not equal to 0
```

#### c) Temp Change ssp245: TPI
```{r Does Temperature scale with TPI?}
explore_cor(x= TPI, y= TempChange_ssp370_2040, 
            x_name= "TPI", 
            y_name= "Temperature Change, Current - 2041 ssp240", 
            cor.method = 'spearman')

# rho = -0.2820622 , p-value < 2.2e-16, alternative hypothesis: true rho is not equal to 0
```

## 3) Explore Precipitation climate change correlations

#### a) Prec Change ssp245: Elevation
```{r Does Temperature scale with elevation?}
explore_cor(x= logElev, y= PrecChange_ssp245_2040, 
            x_name= "log transformed Elevation", 
            y_name= "Precipitation Change, Current - 2041 ssp245", 
            cor.method = 'spearman')
```

#### b) Prec Change ssp245: CTI
```{r Does Precipitation scale with Wetness Index?}
explore_cor(x= CTI, y= PrecChange_ssp245_2040, 
            x_name= "Wetness Index (CTI)", 
            y_name= "Precipitation Change, Current - 2041 ssp245", 
            cor.method = 'spearman')

# rho = -0.2703917, p-value < 2.2e-16, alternative hypothesis: true rho is not equal to 0
```

#### c) Prec Change ssp245: TPI
```{r Does Precipitation scale with TPI?}
explore_cor(x= TPI, y= PrecChange_ssp245_2040, 
            x_name= "TPI", 
            y_name= "Precipitation Change, Current - 2041 ssp240", 
            cor.method = 'spearman')
```

This process can be repeated with the ssp370 data provided. We found in other analysis that 245 better represented the needs of our study, so is displayed here. 