---
title: "Cola_dataprepping"
author: "Kristian Suszczenia"
date: "2024-11-29"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Example with HD

1) Creating Projected Coords:

Before this, I first imported the corrdintaes (PXY_HD.csv, in the Saved2 and Ponly subfolder) to ARCGIS via coordinate conversion, on a WGS 84 map (unprojected), and then exported that as a shape file. These should all be saved in one place, shape files come with multiple files that it links to.

I load that shapefile here
```{r}
#Load unprojected shapefile
HD_unproj_coord <- vect("~/ArcGIS/Projects/ConnectAnalysis_present_unprojected/unprojectedcoords.shp")

#check crs
crs(HD_unproj_coord) #it is in the wrong crs

#So save them in sf package
#use the old projection code just to set it correctly
HD_unproj_cooord_sf <- st_as_sf(HD_unproj_coord, coords = c("x", "y"), crs = 4326) 

#use the sf transform function to ADD coords of the new projection, using the ESRI code for albers conical SEA
HD_coord_proj_sf <- st_transform(HD_unproj_cooord_sf, crs = "ESRI:102028")
crs(HD_coord_proj_sf)

#Extract the new cooridntaes from the geometry column
HD_proj_coords <- st_coordinates(st_geometry(HD_coord_proj_sf))

#Convert the coordinates to a data frame
HD_proj_coords_df <- as.data.frame(HD_proj_coords)


#save
write.csv(HD_proj_coords_df, "~/Directories/R_directories/Masters_VS_borneo/Saved2/Cola_ready/HD/HD_coords_proj.csv", row.names = FALSE)

#Go into this folder, open and check the data is as needed (X and Y headings), then change the extension to .XY

####

#We can also save it as a shapefile by opening this csv in R into an Albers map and using coordinate conversion to save as .shp.
#Save as shp file, this is cola ready too!
st_write(HD_coord_proj_sf, "~/Directories/R_directories/Masters_VS_borneo/Saved2/Cola_ready/HD/HD_coords_proj.shp")
```


2) Creating the projected Raster

This is more simple, all that is needed is the perfect suitability raster from the VS r session, and the rest can be done here in R aside from setting equal dimensions.

```{r}
#library(spatialEco)

#Load the new projected HD HSI raster
HD_perfectHSI <- rast("C:\\Users\\shil5756\\Documents\\Directories\\R_directories\\Masters_VS_borneo\\Saved2\\Saved_realhabitatsuitability\\Suitability_HD1.tif")

#Project it onto our desired Albers projection
HD_HSI_proj <- project(HD_perfectHSI, crs("ESRI:102028"))

#Invert the projected raster linearly
HD_invert_proj <- raster.invert(HD_HSI_proj)

#Then stretch to make the resistance raster
HD_resi_proj <- raster.transformation(HD_invert_proj, trans = "stretch", smin=1, smax= 99)

#make sure the projection has stayed
crs(HD_resi_proj) # it has "PROJCRS[\"Asia_South_Albers_Equal_Area_Conic\",\n 

#Check if x and Y cell dimensions are equal
res(HD_resi_proj) #they seem to be, but in fact are off by a small amount (due to how computers use floating point calculations)

#Thus we need to save them to do the final part in R, the tool 'Resample' (method bilinear) can be used to set the exact resolution (same as cell dimesnions). Load them into a map with the correct Albers project. Then Resample to 925x925.
#save as TIFF to be taken to ARCGIS
writeRaster(HD_resi_proj, 
            "~/ArcGIS/Projects/Cola_dataprep_WGS84/InputRasters/HD_projectedResi.tif")

#Once done, 

#Then we load this bakc into R to restretch 
HD_resi_resample <- rast("~/ArcGIS/Projects/Cola_dataprep_WGS84/ResampledRasters/HD_projectedResi_Resample.tif")
HD_resi_resample_stretch <- raster.transformation(HD_resi_resample, trans = "stretch", smin=1, smax= 99)
plot(HD_resi_resample_stretch)

#and then finally save as an tif, by just using the .tif extension
writeRaster(HD_resi_resample_stretch, "~/Directories/R_directories/Masters_VS_borneo/Saved2/Cola_ready/HD/HD_resi.tif")
#and as asc
writeRaster(HD_resi_resample_stretch, "~/Directories/R_directories/Masters_VS_borneo/Saved2/Cola_ready/HD/HD_resi.ASC")

#however, this gives the asc with NAN nodata values, if you want -9999, you have to open the riff in R and specify it in the tool Raster to ASCII
```

This creates your cola ready folder with the projected coords (csv, XY, and shp. files) and the projected resistance (tif for Cola, and ASC9999 for raw UNICOR)

## 2) HG

1) Creating Projected Coords:

Before this, I first imported the corrdintaes (PXY_HD.csv, in the Saved2 and Ponly subfolder) to ARCGIS via coordinate conversion, on a WGS 84 map (unprojected), and then exported that as a shape file. These should all be saved in one place, shape files come with multiple files that it links to.

I load that shapefile here
```{r}
#Load unprojected shapefile
HG_unproj_coord <- vect("~/ArcGIS/Projects/Cola_dataprep_WGS84/Outputcoords/HWunproj_coord.shp")

#check crs
crs(HG_unproj_coord) #it is in the wrong crs

#So save them in sf package
#use the old projection code just to set it correctly
HG_unproj_cooord_sf <- st_as_sf(HG_unproj_coord, coords = c("x", "y"), crs = 4326) 

#use the sf transform function to ADD coords of the new projection, using the ESRI code for albers conical SEA
HG_coord_proj_sf <- st_transform(HG_unproj_cooord_sf, crs = "ESRI:102028")
crs(HG_coord_proj_sf)

#Extract the new cooridntaes from the geometry column
HG_proj_coords <- st_coordinates(st_geometry(HG_coord_proj_sf))

#Convert the coordinates to a data frame
HG_proj_coords_df <- as.data.frame(HG_proj_coords)

#save
write.csv(HG_proj_coords_df, "~/Directories/R_directories/Masters_VS_borneo/Saved2/Cola_ready/HG/HG_coords_proj.csv", row.names = FALSE)

#Go into this folder, open and check the data is as needed (X and Y headings), then change the extension to .XY

####

#We can also save it as a shapefile by opening this csv in R into an Albers map and using coordinate conversion to save as .shp.
#Save as shp file, this is cola ready too!
st_write(HG_coord_proj_sf, "~/Directories/R_directories/Masters_VS_borneo/Saved2/Cola_ready/HG/HG_coords_proj.shp")
```

2) Creating the projected Raster

```{r}
#library(spatialEco)

#Load the new projected HD HSI raster
HG_perfectHSI <- rast("C:\\Users\\shil5756\\Documents\\Directories\\R_directories\\Masters_VS_borneo\\Saved2\\Saved_realhabitatsuitability\\Suitability_HG1.tif")

#Project it onto our desired Albers projection
HG_HSI_proj <- project(HG_perfectHSI, crs("ESRI:102028"))

#Invert the projected raster linearly
HG_invert_proj <- raster.invert(HG_HSI_proj)

#Then stretch to make the resistance raster
HG_resi_proj <- raster.transformation(HG_invert_proj, trans = "stretch", smin=1, smax= 99)

#make sure the projection has stayed
crs(HG_resi_proj) # it has "PROJCRS[\"Asia_South_Albers_Equal_Area_Conic\",\n 

#Check if x and Y cell dimensions are equal
res(HG_resi_proj) #they seem to be, but in fact are off by a small amount (due to how computers use floating point calculations)

#Thus we need to save them to do the final part in R, the tool 'Resample' (method bilinear) can be used to set the exact resolution (same as cell dimesnions). Load them into a map with the correct Albers project. Then Resample to 925x925.
#save as TIFF to be taken to ARCGIS
writeRaster(HG_resi_proj, 
            "~/ArcGIS/Projects/Cola_dataprep_WGS84/InputRasters/HG_projectedResi.tif")

#Once done, 

#Then we load this bakc into R to restretch 
HG_resi_resample <- rast("~/ArcGIS/Projects/Cola_dataprep_WGS84/ResampledRasters/HG_projectedResi_Resample.tif")
HG_resi_resample_stretch <- raster.transformation(HG_resi_resample, trans = "stretch", smin=1, smax= 99)
plot(HG_resi_resample_stretch)

#and then finally save as an tif, by just using the .tif extension
writeRaster(HG_resi_resample_stretch, "~/Directories/R_directories/Masters_VS_borneo/Saved2/Cola_ready/HG/HG_resi.tif")
#and as asc
writeRaster(HG_resi_resample_stretch, "~/Directories/R_directories/Masters_VS_borneo/Saved2/Cola_ready/HG/HG_resi.ASC")

#however, this gives the asc with NAN nodata values, if you want -9999, you have to open the riff in R and specify it in the tool Raster to ASCII
```

This creates your cola ready folder with the projected coords (csv, XY, and shp. files) and the projected resistance (tif for Cola, and ASC9999 for raw UNICOR)


## 3) HW

1) Creating Projected Coords:

Before this, I first imported the corrdintaes (PXY_HD.csv, in the Saved2 and Ponly subfolder) to ARCGIS via coordinate conversion, on a WGS 84 map (unprojected), and then exported that as a shape file. These should all be saved in one place, shape files come with multiple files that it links to.

I load that shapefile here
```{r}
#Load unprojected shapefile
HW_unproj_coord <- vect("~/ArcGIS/Projects/Cola_dataprep_WGS84/Outputcoords/HWunproj_coord.shp")

#check crs
crs(HW_unproj_coord) #it is in the wrong crs

#So save them in sf package
#use the old projection code just to set it correctly
HW_unproj_cooord_sf <- st_as_sf(HW_unproj_coord, coords = c("x", "y"), crs = 4326) 

#use the sf transform function to ADD coords of the new projection, using the ESRI code for albers conical SEA
HW_coord_proj_sf <- st_transform(HW_unproj_cooord_sf, crs = "ESRI:102028")
crs(HW_coord_proj_sf)

#Extract the new cooridntaes from the geometry column
HW_proj_coords <- st_coordinates(st_geometry(HW_coord_proj_sf))

#Convert the coordinates to a data frame
HW_proj_coords_df <- as.data.frame(HW_proj_coords)

#save
write.csv(HW_proj_coords_df, "~/Directories/R_directories/Masters_VS_borneo/Saved2/Cola_ready/HW/HW_coords_proj.csv", row.names = FALSE)

#Go into this folder, open and check the data is as needed (X and Y headings), then change the extension to .XY

####

#We can also save it as a shapefile by opening this csv in R into an Albers map and using coordinate conversion to save as .shp.
#Save as shp file, this is cola ready too!
st_write(HW_coord_proj_sf, "~/Directories/R_directories/Masters_VS_borneo/Saved2/Cola_ready/HW/HW_coords_proj.shp")
```

2) Creating the projected Raster

```{r}
#library(spatialEco)

#Load the new projected HD HSI raster
HW_perfectHSI <- rast("C:\\Users\\shil5756\\Documents\\Directories\\R_directories\\Masters_VS_borneo\\Saved2\\Saved_realhabitatsuitability\\Suitability_HW1.tif")

#Project it onto our desired Albers projection
HW_HSI_proj <- project(HW_perfectHSI, crs("ESRI:102028"))

#Invert the projected raster linearly
HW_invert_proj <- raster.invert(HW_HSI_proj)

#Then stretch to make the resistance raster
HW_resi_proj <- raster.transformation(HW_invert_proj, trans = "stretch", smin=1, smax= 99)

#make sure the projection has stayed
crs(HW_resi_proj) # it has "PROJCRS[\"Asia_South_Albers_Equal_Area_Conic\",\n 

#Check if x and Y cell dimensions are equal
res(HW_resi_proj) #they seem to be, but in fact are off by a small amount (due to how computers use floating point calculations)

#Thus we need to save them to do the final part in R, the tool 'Resample' (method bilinear) can be used to set the exact resolution (same as cell dimesnions). Load them into a map with the correct Albers project. Then Resample to 925x925.
#save as TIFF to be taken to ARCGIS
writeRaster(HW_resi_proj, 
            "~/ArcGIS/Projects/Cola_dataprep_WGS84/InputRasters/HW_projectedResi.tif")

#Once done, 

#Then we load this bakc into R to restretch 
HW_resi_resample <- rast("~/ArcGIS/Projects/Cola_dataprep_WGS84/ResampledRasters/HW_projectedResi_Resample.tif")
HW_resi_resample_stretch <- raster.transformation(HW_resi_resample, trans = "stretch", smin=1, smax= 99)
plot(HW_resi_resample_stretch)

#and then finally save as an tif, by just using the .tif extension
writeRaster(HW_resi_resample_stretch, "~/Directories/R_directories/Masters_VS_borneo/Saved2/Cola_ready/HW/HW_resi.tif")
#and as asc
writeRaster(HW_resi_resample_stretch, "~/Directories/R_directories/Masters_VS_borneo/Saved2/Cola_ready/HW/HW_resi.ASC")

#however, this gives the asc with NAN nodata values, if you want -9999, you have to open the riff in R and specify it in the tool Raster to ASCII
```

This creates your cola ready folder with the projected coords (csv, XY, and shp. files) and the projected resistance (tif for Cola, and ASC9999 for raw UNICOR)

## 4) LD

1) Creating Projected Coords:

Before this, I first imported the corrdintaes (PXY_HD.csv, in the Saved2 and Ponly subfolder) to ARCGIS via coordinate conversion, on a WGS 84 map (unprojected), and then exported that as a shape file. These should all be saved in one place, shape files come with multiple files that it links to.

I load that shapefile here
```{r}
#Load unprojected shapefile
LD_unproj_coord <- vect("~/ArcGIS/Projects/Cola_dataprep_WGS84/Outputcoords/LDunproj_coord.shp")

#check crs
crs(LD_unproj_coord) #it is in the wrong crs

#So save them in sf package
#use the old projection code just to set it correctly
LD_unproj_cooord_sf <- st_as_sf(LD_unproj_coord, coords = c("x", "y"), crs = 4326) 

#use the sf transform function to ADD coords of the new projection, using the ESRI code for albers conical SEA
LD_coord_proj_sf <- st_transform(LD_unproj_cooord_sf, crs = "ESRI:102028")
crs(LD_coord_proj_sf)

#Extract the new cooridntaes from the geometry column
LD_proj_coords <- st_coordinates(st_geometry(LD_coord_proj_sf))

#Convert the coordinates to a data frame
LD_proj_coords_df <- as.data.frame(LD_proj_coords)

#save
write.csv(LD_proj_coords_df, "~/Directories/R_directories/Masters_VS_borneo/Saved2/Cola_ready/LD/LD_coords_proj.csv", row.names = FALSE)

#Go into this folder, open and check the data is as needed (X and Y headings), then change the extension to .XY

####

#We can also save it as a shapefile by opening this csv in R into an Albers map and using coordinate conversion to save as .shp.
#Save as shp file, this is cola ready too!
st_write(LD_coord_proj_sf, "~/Directories/R_directories/Masters_VS_borneo/Saved2/Cola_ready/LD/LD_coords_proj.shp")
```

2) Creating the projected Raster

```{r}
#library(spatialEco)

#Load the new projected HD HSI raster
LD_perfectHSI <- rast("C:\\Users\\shil5756\\Documents\\Directories\\R_directories\\Masters_VS_borneo\\Saved2\\Saved_realhabitatsuitability\\Suitability_LD1.tif")

#Project it onto our desired Albers projection
LD_HSI_proj <- project(LD_perfectHSI, crs("ESRI:102028"))

#Invert the projected raster linearly
LD_invert_proj <- raster.invert(LD_HSI_proj)

#Then stretch to make the resistance raster
LD_resi_proj <- raster.transformation(LD_invert_proj, trans = "stretch", smin=1, smax= 99)

#make sure the projection has stayed
crs(LD_resi_proj) # it has "PROJCRS[\"Asia_South_Albers_Equal_Area_Conic\",\n 

#Check if x and Y cell dimensions are equal
res(LD_resi_proj) #they seem to be, but in fact are off by a small amount (due to how computers use floating point calculations)

#Thus we need to save them to do the final part in R, the tool 'Resample' (method bilinear) can be used to set the exact resolution (same as cell dimesnions). Load them into a map with the correct Albers project. Then Resample to 925x925.
#save as TIFF to be taken to ARCGIS
writeRaster(LD_resi_proj, 
            "~/ArcGIS/Projects/Cola_dataprep_WGS84/InputRasters/LD_projectedResi.tif")

#Once done, 

#Then we load this bakc into R to restretch 
LD_resi_resample <- rast("~/ArcGIS/Projects/Cola_dataprep_WGS84/ResampledRasters/LD_projectedResi_Resample.tif")
LD_resi_resample_stretch <- raster.transformation(LD_resi_resample, trans = "stretch", smin=1, smax= 99)
plot(LD_resi_resample_stretch)

#and then finally save as an tif, by just using the .tif extension
writeRaster(LD_resi_resample_stretch, "~/Directories/R_directories/Masters_VS_borneo/Saved2/Cola_ready/LD/LD_resi.tif")
#and as asc
writeRaster(LD_resi_resample_stretch, "~/Directories/R_directories/Masters_VS_borneo/Saved2/Cola_ready/LD/LD_resi.ASC")

#however, this gives the asc with NAN nodata values, if you want -9999, you have to open the riff in R and specify it in the tool Raster to ASCII
```

This creates your cola ready folder with the projected coords (csv, XY, and shp. files) and the projected resistance (tif for Cola, and ASC9999 for raw UNICOR)

## 5) LG

1) Creating Projected Coords:

Before this, I first imported the corrdintaes (PXY_HD.csv, in the Saved2 and Ponly subfolder) to ARCGIS via coordinate conversion, on a WGS 84 map (unprojected), and then exported that as a shape file. These should all be saved in one place, shape files come with multiple files that it links to.

I load that shapefile here
```{r}
#Load unprojected shapefile
LG_unproj_coord <- vect("~/ArcGIS/Projects/Cola_dataprep_WGS84/Outputcoords/LGunproj_coord.shp")

#check crs
crs(LG_unproj_coord) #it is in the wrong crs

#So save them in sf package
#use the old projection code just to set it correctly
LG_unproj_cooord_sf <- st_as_sf(LG_unproj_coord, coords = c("x", "y"), crs = 4326) 

#use the sf transform function to ADD coords of the new projection, using the ESRI code for albers conical SEA
LG_coord_proj_sf <- st_transform(LG_unproj_cooord_sf, crs = "ESRI:102028")
crs(LG_coord_proj_sf)

#Extract the new cooridntaes from the geometry column
LG_proj_coords <- st_coordinates(st_geometry(LG_coord_proj_sf))

#Convert the coordinates to a data frame
LG_proj_coords_df <- as.data.frame(LG_proj_coords)

#save
write.csv(LG_proj_coords_df, "~/Directories/R_directories/Masters_VS_borneo/Saved2/Cola_ready/LG/LG_coords_proj.csv", row.names = FALSE)

#Go into this folder, open and check the data is as needed (X and Y headings), then change the extension to .XY

####

#We can also save it as a shapefile by opening this csv in R into an Albers map and using coordinate conversion to save as .shp.
#Save as shp file, this is cola ready too!
st_write(LG_coord_proj_sf, "~/Directories/R_directories/Masters_VS_borneo/Saved2/Cola_ready/LG/LG_coords_proj.shp")
```

2) Creating the projected Raster

```{r}
#library(spatialEco)

#Load the new projected HD HSI raster
LG_perfectHSI <- rast("C:\\Users\\shil5756\\Documents\\Directories\\R_directories\\Masters_VS_borneo\\Saved2\\Saved_realhabitatsuitability\\Suitability_LG1.tif")

#Project it onto our desired Albers projection
LG_HSI_proj <- project(LG_perfectHSI, crs("ESRI:102028"))

#Invert the projected raster linearly
LG_invert_proj <- raster.invert(LG_HSI_proj)

#Then stretch to make the resistance raster
LG_resi_proj <- raster.transformation(LG_invert_proj, trans = "stretch", smin=1, smax= 99)

#make sure the projection has stayed
crs(LG_resi_proj) # it has "PROJCRS[\"Asia_South_Albers_Equal_Area_Conic\",\n 

#Check if x and Y cell dimensions are equal
res(LG_resi_proj) #they seem to be, but in fact are off by a small amount (due to how computers use floating point calculations)

#Thus we need to save them to do the final part in R, the tool 'Resample' (method bilinear) can be used to set the exact resolution (same as cell dimesnions). Load them into a map with the correct Albers project. Then Resample to 925x925.
#save as TIFF to be taken to ARCGIS
writeRaster(LG_resi_proj, 
            "~/ArcGIS/Projects/Cola_dataprep_WGS84/InputRasters/LG_projectedResi.tif")

#Once done, 

#Then we load this bakc into R to restretch 
LG_resi_resample <- rast("~/ArcGIS/Projects/Cola_dataprep_WGS84/ResampledRasters/LG_projectedResi_Resample.tif")
LG_resi_resample_stretch <- raster.transformation(LG_resi_resample, trans = "stretch", smin=1, smax= 99)
plot(LG_resi_resample_stretch)

#and then finally save as an tif, by just using the .tif extension
writeRaster(LG_resi_resample_stretch, "~/Directories/R_directories/Masters_VS_borneo/Saved2/Cola_ready/LG/LG_resi.tif")
#and as asc
writeRaster(LG_resi_resample_stretch, "~/Directories/R_directories/Masters_VS_borneo/Saved2/Cola_ready/LG/LG_resi.ASC")

#however, this gives the asc with NAN nodata values, if you want -9999, you have to open the riff in R and specify it in the tool Raster to ASCII
```

This creates your cola ready folder with the projected coords (csv, XY, and shp. files) and the projected resistance (tif for Cola, and ASC9999 for raw UNICOR)

## 6) LW

1) Creating Projected Coords:

Before this, I first imported the corrdintaes (PXY_HD.csv, in the Saved2 and Ponly subfolder) to ARCGIS via coordinate conversion, on a WGS 84 map (unprojected), and then exported that as a shape file. These should all be saved in one place, shape files come with multiple files that it links to.

I load that shapefile here
```{r}
#Load unprojected shapefile
LW_unproj_coord <- vect("~/ArcGIS/Projects/Cola_dataprep_WGS84/Outputcoords/LWunproj_coord.shp")

#check crs
crs(LW_unproj_coord) #it is in the wrong crs

#So save them in sf package
#use the old projection code just to set it correctly
LW_unproj_cooord_sf <- st_as_sf(LW_unproj_coord, coords = c("x", "y"), crs = 4326) 

#use the sf transform function to ADD coords of the new projection, using the ESRI code for albers conical SEA
LW_coord_proj_sf <- st_transform(LW_unproj_cooord_sf, crs = "ESRI:102028")
crs(LW_coord_proj_sf)

#Extract the new cooridntaes from the geometry column
LW_proj_coords <- st_coordinates(st_geometry(LW_coord_proj_sf))

#Convert the coordinates to a data frame
LW_proj_coords_df <- as.data.frame(LW_proj_coords)

#save
write.csv(LW_proj_coords_df, "~/Directories/R_directories/Masters_VS_borneo/Saved2/Cola_ready/LW/LW_coords_proj.csv", row.names = FALSE)

#Go into this folder, open and check the data is as needed (X and Y headings), then change the extension to .XY

####

#We can also save it as a shapefile by opening this csv in R into an Albers map and using coordinate conversion to save as .shp.
#Save as shp file, this is cola ready too!
st_write(LW_coord_proj_sf, "~/Directories/R_directories/Masters_VS_borneo/Saved2/Cola_ready/LW/LW_coords_proj.shp")
```

2) Creating the projected Raster

```{r}
#library(spatialEco)

#Load the new projected HD HSI raster
LW_perfectHSI <- rast("C:\\Users\\shil5756\\Documents\\Directories\\R_directories\\Masters_VS_borneo\\Saved2\\Saved_realhabitatsuitability\\Suitability_LW1.tif")

#Project it onto our desired Albers projection
LW_HSI_proj <- project(LW_perfectHSI, crs("ESRI:102028"))

#Invert the projected raster linearly
LW_invert_proj <- raster.invert(LW_HSI_proj)

#Then stretch to make the resistance raster
LW_resi_proj <- raster.transformation(LW_invert_proj, trans = "stretch", smin=1, smax= 99)

#make sure the projection has stayed
crs(LW_resi_proj) # it has "PROJCRS[\"Asia_South_Albers_Equal_Area_Conic\",\n 

#Check if x and Y cell dimensions are equal
res(LW_resi_proj) #they seem to be, but in fact are off by a small amount (due to how computers use floating point calculations)

#Thus we need to save them to do the final part in R, the tool 'Resample' (method bilinear) can be used to set the exact resolution (same as cell dimesnions). Load them into a map with the correct Albers project. Then Resample to 925x925.
#save as TIFF to be taken to ARCGIS
writeRaster(LW_resi_proj, 
            "~/ArcGIS/Projects/Cola_dataprep_WGS84/InputRasters/LW_projectedResi.tif")

#Once done, 

#Then we load this bakc into R to restretch 
LW_resi_resample <- rast("~/ArcGIS/Projects/Cola_dataprep_WGS84/ResampledRasters/LW_projectedResi_Resample.tif")
LW_resi_resample_stretch <- raster.transformation(LW_resi_resample, trans = "stretch", smin=1, smax= 99)
plot(LW_resi_resample_stretch)

#and then finally save as an tif, by just using the .tif extension
writeRaster(LW_resi_resample_stretch, "~/Directories/R_directories/Masters_VS_borneo/Saved2/Cola_ready/LW/LW_resi.tif")
#and as asc
writeRaster(LW_resi_resample_stretch, "~/Directories/R_directories/Masters_VS_borneo/Saved2/Cola_ready/LW/LW_resi.ASC")

#however, this gives the asc with NAN nodata values, if you want -9999, you have to open the riff in R and specify it in the tool Raster to ASCII
```

This creates your cola ready folder with the projected coords (csv, XY, and shp. files) and the projected resistance (tif for Cola, and ASC9999 for raw UNICOR)

### Include template for Cola run and link to their guide














