---
title: "SensitivityExploration"
author: "Kristian Suszczenia"
date: "2025-01-20"
output: html_document
---

This code entails the sensitivty analysis and graphs produced in the supplment of our study. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Load required packages
```{r}
library(terra)
library(ggplot2)
library(gridExtra)
library(ragg)
library(here)
```

#functions
```{r}
save_figure <- function(Figure_to_save,
                                  filename, width, height, res, scaling){
  agg_png(filename, 
          width   =  width, 
          height  =  height, 
          units   =  "cm", 
          res     =  res, 
          scaling =  scaling)
  Figure_to_save
  print(Figure_to_save)
  dev.off()
}
```

##Import the band 1 metric to test
I am importing as tif

```{r}
band1TIFF <- rast("C:\\Users\\shil5756\\Documents\\ConnectivityData_master\\SensitivityAnalysis\\MetricOutputTIFs\\Band1.tif")

#you can look at the object
band1TIFF
#get sumary stats
summary(band1TIFF)
#and plot it easily!
plot(band1TIFF)
```

But, mainly, we want to save the mean, median, and 2SD
```{r}
band1values <- na.omit(values(band1TIFF))
band1mean <- mean(band1values) # = 1.463686
band1median <- median(band1values) # = 0
band1SD <- sd(band1values) # = 6.831418

#and add the computation time
band1time <- 40

```

#Repeat to get values for all the bands

2 band
```{r}
band2TIFF <- rast("C:\\Users\\shil5756\\Documents\\ConnectivityData_master\\SensitivityAnalysis\\MetricOutputTIFs\\band2.tif")
band2values <- na.omit(values(band2TIFF))
band2mean <- mean(band2values) # = 1.970988
band2median <- median(band2values) # = 0
band2SD <- sd(band2values) # = 7.061939
band2time <- 70
```

3 band
```{r}
band3TIFF <- rast("C:\\Users\\shil5756\\Documents\\ConnectivityData_master\\SensitivityAnalysis\\MetricOutputTIFs\\band3.tif")
band3values <- na.omit(values(band3TIFF))
band3mean <- mean(band3values) # = 4.8856
band3median <- median(band3values) # = 0.7585
band3SD <- sd(band3values) # = 12.9285
band3time <- 77
```

4band
```{r}
band4TIFF <- rast("C:\\Users\\shil5756\\Documents\\ConnectivityData_master\\SensitivityAnalysis\\MetricOutputTIFs\\band4.tif")
band4values <- na.omit(values(band4TIFF))
band4mean <- mean(band4values) # = 6.3546
band4median <- median(band4values) # = 1.4470
band4SD <- sd(band4values) # = 14.1165
band4time <- 97
```

5band
```{r}
band5TIFF <- rast("C:\\Users\\shil5756\\Documents\\ConnectivityData_master\\SensitivityAnalysis\\MetricOutputTIFs\\band5.tif")
band5values <- na.omit(values(band5TIFF))
band5mean <- mean(band5values) 
band5median <- median(band5values) 
band5SD <- sd(band5values)
band5time <- 119
```

10band
```{r}
band10TIFF <- rast("C:\\Users\\shil5756\\Documents\\ConnectivityData_master\\SensitivityAnalysis\\MetricOutputTIFs\\band10.tif")
band10values <- na.omit(values(band10TIFF))
band10mean <- mean(band10values) 
band10median <- median(band10values) 
band10SD <- sd(band10values) 
band10time <- 177
```

20band
```{r}
band20TIFF <- rast("C:\\Users\\shil5756\\Documents\\ConnectivityData_master\\SensitivityAnalysis\\MetricOutputTIFs\\20band.tif")
band20values <- na.omit(values(band20TIFF))
band20mean <- mean(band20values) 
band20median <- median(band20values) 
band20SD <- sd(band20values) 
band20time <- 370
```

exponential 4 band
```{r}
expband4TIFF <- rast("C:\\Users\\shil5756\\Documents\\ConnectivityData_master\\SensitivityAnalysis\\MetricOutputTIFs\\exp4band.tif")
expband4values <- na.omit(values(expband4TIFF))
expband4mean <- mean(expband4values) 
expband4median <- median(expband4values) 
expband4SD <- sd(expband4values) 
expband4time <- 99
```

exponential 5 band
```{r}
expband5TIFF <- rast("C:\\Users\\shil5756\\Documents\\ConnectivityData_master\\SensitivityAnalysis\\MetricOutputTIFs\\exp5band.tif")
expband5values <- na.omit(values(expband5TIFF))
expband5mean <- mean(expband5values) 
expband5median <- median(expband5values) 
expband5SD <- sd(expband5values) 
expband5time <- 138
```

exponential 6 band
```{r}
expband6TIFF <- rast("C:\\Users\\shil5756\\Documents\\ConnectivityData_master\\SensitivityAnalysis\\MetricOutputTIFs\\exp6band.tif")
expband6values <- na.omit(values(expband6TIFF))
expband6mean <- mean(expband6values) 
expband6median <- median(expband6values) 
expband6SD <- sd(expband6values) 
expband6time <- 148
```

#collate the data into a single object so we can plot it
```{r}
#Build each column
names <- c('1band', '2band', '3band', '4band', '5band', '10band', '20band', 'exp4band', 'exp5band', 'exp6band')
means <- c(band1mean, band2mean, band3mean, band4mean, band5mean, band10mean, band20mean, expband4mean, expband5mean, expband6mean)
medians <- c(band1median, band2median, band3median, band4median, band5median, band10median, band20median, expband4median, expband5median, expband6median)
bandnum <- c(1, 2, 3, 4, 5, 10, 20, 4, 5, 6)
comptime <- c(band1time, band2time, band3time, band4time, band5time, band10time, band20time, expband4time, expband5time, expband6time)
metrictype <- c('linear', 'linear', 'linear', 'linear', 'linear', 'linear', 'linear', 'exponential', 'exponential', 'exponential')

#Build the dataframe 
BandStatsDf <- data.frame(Bandname= names, mean= means, median= medians, bands= bandnum, computation_time= comptime, metric_type= metrictype)

#Set the catagorical variables as catagorical (as 'factor')
BandStatsDf$metric_type <- factor(BandStatsDf$metric_type, levels = c('exponential', 'linear'))
```

plot the means against number of bands
```{r}
ggplot(BandStatsDf, 
       aes(x = bands, 
           y = mean, 
           colour = metric_type)) +
  geom_point(size = 3) + 
  labs(title = "Sensitivty Analysis of Number of strength bands", 
       x = "Number of Bands", 
       y = "Mean Vulnerability Estimate") +
  theme_linedraw()

#save it
agg_png("C:/Users/shil5756/Pictures/Saved Pictures/SensitivityAnalysisR/Mean_V_Bandnumber.png", width = 3000, height = 2000, res = 300)

ggplot(BandStatsDf, 
       aes(x = bands, 
           y = mean, 
           colour = metric_type)) +
  geom_point(size = 3) + 
  labs(title = "Sensitivty Analysis of Number of strength bands", 
       x = "Number of Bands", 
       y = "Mean Vulnerability Estimate") +
  theme_linedraw()

# Close the PNG device to save the file
dev.off()
```

plot the medians against number of bands
```{r}
ggplot(BandStatsDf, 
       aes(x = bands, 
           y = median, 
           colour = metric_type)) +
  geom_point(size = 3) + 
  labs(title = "Sensitivty Analysis of Number of strength bands", 
       x = "Number of Bands", 
       y = "Median Vulnerability Estimate") +
  theme_minimal()

#save it
agg_png("C:/Users/shil5756/Pictures/Saved Pictures/SensitivityAnalysisR/Median_V_Bandnumber.png", width = 3000, height = 2000, res = 300)

ggplot(BandStatsDf, 
       aes(x = bands, 
           y = median, 
           colour = metric_type)) +
  geom_point(size = 3) + 
  labs(title = "Sensitivty Analysis of Number of strength bands", 
       x = "Number of Bands", 
       y = "Median Vulnerability Estimate") +
  theme_linedraw()

# Close the PNG device to save the file
dev.off()
```

plot the computation time against number of bands
```{r}
ggplot(BandStatsDf, 
       aes(x = bands, 
           y = computation_time, 
           colour = metric_type)) +
  geom_point(size = 3) + 
  labs(title = "Sensitivty Analysis of Number of strength bands", 
       x = "Number of Bands", 
       y = "Computation Time") +
  theme_minimal()

#save it
agg_png("C:/Users/shil5756/Pictures/Saved Pictures/SensitivityAnalysisR/CompTime_V_Bandnumber.png", width = 3000, height = 2000, res = 300)

ggplot(BandStatsDf, 
       aes(x = bands, 
           y = computation_time, 
           colour = metric_type)) +
  geom_point(size = 3) + 
  labs(title = "Sensitivty Analysis of Number of strength bands", 
       x = "Number of Bands", 
       y = "Computation Time") +
  theme_linedraw()

# Close the PNG device to save the file
dev.off()
```

plot the computation time against means
```{r}
ggplot(BandStatsDf, 
       aes(x = mean, 
           y = computation_time, 
           colour = metric_type,
           size = bands)) +
  geom_point() + 
  labs(title = "Sensitivty Analysis of Number of strength bands", 
       x = "Mean Vulnerability Estimate", 
       y = "Computation Time") +
  theme_minimal()
```

#Explore Histogram

1 Band
```{r}
summary(band1values)

ggplot(band1values,
       aes(x= Band_1))+
  geom_histogram(bins = 100)+
  geom_vline(aes(xintercept = median(Band_1)), color = "red", linetype = "dashed", size = 0.75)+
  geom_vline(aes(xintercept = quantile(Band_1, 0.25)), color = "green", linetype = "dotted", size = 1) +
  geom_vline(aes(xintercept = quantile(Band_1, 0.75)), color = "blue", linetype = "dotted", size = 1) +
  labs(title = "1 Band", 
       x = "Vulnerability", 
       y = "Frequency")
```

2 Band
```{r}
summary(band2values)

ggplot(band2values,
       aes(x= Band_1))+
  geom_histogram(bins = 100)+
  geom_vline(aes(xintercept = median(Band_1)), color = "red", linetype = "dashed", size = 0.75)+
  geom_vline(aes(xintercept = quantile(Band_1, 0.25)), color = "green", linetype = "dotted", size = 1) +
  geom_vline(aes(xintercept = quantile(Band_1, 0.75)), color = "blue", linetype = "dotted", size = 1) +
  labs(title = "2 Band", 
       x = "Vulnerability", 
       y = "Frequency")
```

3 Band
```{r}
summary(band3values)

ggplot(band3values,
       aes(x= Band_1))+
  geom_histogram(bins = 100)+
  geom_vline(aes(xintercept = median(Band_1)), color = "red", linetype = "dashed", size = 0.75)+
  geom_vline(aes(xintercept = quantile(Band_1, 0.25)), color = "green", linetype = "dotted", size = 1) +
  geom_vline(aes(xintercept = quantile(Band_1, 0.75)), color = "blue", linetype = "dotted", size = 1) +
  labs(title = "3 Band", 
       x = "Vulnerability", 
       y = "Frequency")
```

4 Band
```{r}
summary(band4values)

ggplot(band4values,
       aes(x= Band_1))+
  geom_histogram(bins = 100)+
  geom_vline(aes(xintercept = median(Band_1)), color = "red", linetype = "dashed", size = 0.75)+
  geom_vline(aes(xintercept = quantile(Band_1, 0.25)), color = "green", linetype = "dotted", size = 1) +
  geom_vline(aes(xintercept = quantile(Band_1, 0.75)), color = "blue", linetype = "dotted", size = 1) +
  labs(title = "4 Band", 
       x = "Vulnerability", 
       y = "Frequency")
```

5 Band
```{r}
summary(band5values)

ggplot(band5values,
       aes(x= Band_1))+
  geom_histogram(bins = 100)+
  geom_vline(aes(xintercept = median(Band_1)), color = "red", linetype = "dashed", size = 0.75)+
  geom_vline(aes(xintercept = quantile(Band_1, 0.25)), color = "green", linetype = "dotted", size = 1) +
  geom_vline(aes(xintercept = quantile(Band_1, 0.75)), color = "blue", linetype = "dotted", size = 1) +
  labs(title = "5 Band", 
       x = "Vulnerability", 
       y = "Frequency")
```

10 Band
```{r}
summary(band10values)

ggplot(band10values,
       aes(x= Band_1))+
  geom_histogram(bins = 100)+
  geom_vline(aes(xintercept = median(Band_1)), color = "red", linetype = "dashed", size = 0.75)+
  geom_vline(aes(xintercept = quantile(Band_1, 0.25)), color = "green", linetype = "dotted", size = 1) +
  geom_vline(aes(xintercept = quantile(Band_1, 0.75)), color = "blue", linetype = "dotted", size = 1) +
  labs(title = "10 Band", 
       x = "Vulnerability", 
       y = "Frequency")
```

20 Band
```{r}
summary(band20values)

ggplot(band20values,
       aes(x= Band_1))+
  geom_histogram(bins = 100)+
  geom_vline(aes(xintercept = median(Band_1)), color = "red", linetype = "dashed", size = 0.75)+
  geom_vline(aes(xintercept = quantile(Band_1, 0.25)), color = "green", linetype = "dotted", size = 1) +
  geom_vline(aes(xintercept = quantile(Band_1, 0.75)), color = "blue", linetype = "dotted", size = 1) +
  labs(title = "20 Band", 
       x = "Vulnerability", 
       y = "Frequency")
```

Exponential 4 Band
```{r}
summary(expband4values)

ggplot(expband4values,
       aes(x= Band_1))+
  geom_histogram(bins = 100)+
  geom_vline(aes(xintercept = median(Band_1)), color = "red", linetype = "dashed", size = 0.75)+
  geom_vline(aes(xintercept = quantile(Band_1, 0.25)), color = "green", linetype = "dotted", size = 1) +
  geom_vline(aes(xintercept = quantile(Band_1, 0.75)), color = "blue", linetype = "dotted", size = 1) +
  labs(title = "Exponential 4 Band", 
       x = "Vulnerability", 
       y = "Frequency")
```

Exponential 5 Band
```{r}
summary(expband5values)

ggplot(expband5values,
       aes(x= Band_1))+
  geom_histogram(bins = 100)+
  geom_vline(aes(xintercept = median(Band_1)), color = "red", linetype = "dashed", size = 0.75)+
  geom_vline(aes(xintercept = quantile(Band_1, 0.25)), color = "green", linetype = "dotted", size = 1) +
  geom_vline(aes(xintercept = quantile(Band_1, 0.75)), color = "blue", linetype = "dotted", size = 1) +
  labs(title = "Exponential 5 Band", 
       x = "Vulnerability", 
       y = "Frequency")
```

Exponential 6 Band
```{r}
summary(expband6values)

ggplot(expband6values,
       aes(x= Band_1))+
  geom_histogram(bins = 100)+
  geom_vline(aes(xintercept = median(Band_1)), color = "red", linetype = "dashed", size = 0.75)+
  geom_vline(aes(xintercept = quantile(Band_1, 0.25)), color = "green", linetype = "dotted", size = 1) +
  geom_vline(aes(xintercept = quantile(Band_1, 0.75)), color = "blue", linetype = "dotted", size = 1) +
  labs(title = "Exponential 6 Band", 
       x = "Vulnerability", 
       y = "Frequency")
```

## Experiment with transformations

```{r}
# Log transformation (log(x) or log(x + 1) to handle zeros)
ExpBand5_logvalues <- log(expband5values + 1)  # Use log(x + 1) to avoid log(0)

ggplot(ExpBand5_logvalues,
       aes(x= Band_1))+
  geom_histogram(bins = 100)+
  geom_vline(aes(xintercept = median(Band_1)), color = "red", linetype = "dashed", size = 0.75)+
  geom_vline(aes(xintercept = quantile(Band_1, 0.25)), color = "green", linetype = "dotted", size = 1) +
  geom_vline(aes(xintercept = quantile(Band_1, 0.75)), color = "blue", linetype = "dotted", size = 1) +
  labs(title = "Exponential 6 Band", 
       x = "LOG Vulnerability", 
       y = "Frequency")
```

```{r}
# Log transformation (log(x) or log(x + 1) to handle zeros)
ExpBand6_logvalues <- log(expband6values + 1)  # Use log(x + 1) to avoid log(0)

ggplot(ExpBand6_logvalues,
       aes(x= Band_1))+
  geom_histogram(bins = 100)+
  geom_vline(aes(xintercept = median(Band_1)), color = "red", linetype = "dashed", size = 0.75)+
  geom_vline(aes(xintercept = quantile(Band_1, 0.25)), color = "green", linetype = "dotted", size = 1) +
  geom_vline(aes(xintercept = quantile(Band_1, 0.75)), color = "blue", linetype = "dotted", size = 1) +
  labs(title = "Exponential 6 Band", 
       x = "LOG Vulnerability", 
       y = "Frequency")
```

```{r}
# Log transformation (log(x) or log(x + 1) to handle zeros)
Band10_logvalues <- log(band10values + 1)  # Use log(x + 1) to avoid log(0)

ggplot(Band10_logvalues,
       aes(x= Band_1))+
  geom_histogram(bins = 100)+
  geom_vline(aes(xintercept = median(Band_1)), color = "red", linetype = "dashed", size = 0.75)+
  geom_vline(aes(xintercept = quantile(Band_1, 0.25)), color = "green", linetype = "dotted", size = 1) +
  geom_vline(aes(xintercept = quantile(Band_1, 0.75)), color = "blue", linetype = "dotted", size = 1) +
  labs(title = "20 Band", 
       x = "LOG Vulnerability", 
       y = "Frequency")
```

```{r}
# Log transformation (log(x) or log(x + 1) to handle zeros)
Band20_logvalues <- log(band20values + 1)  # Use log(x + 1) to avoid log(0)

ggplot(Band20_logvalues,
       aes(x= Band_1))+
  geom_histogram(bins = 100)+
  geom_vline(aes(xintercept = median(Band_1)), color = "red", linetype = "dashed", size = 0.75)+
  geom_vline(aes(xintercept = quantile(Band_1, 0.25)), color = "green", linetype = "dotted", size = 1) +
  geom_vline(aes(xintercept = quantile(Band_1, 0.75)), color = "blue", linetype = "dotted", size = 1) +
  labs(title = "20 Band", 
       x = "LOG Vulnerability", 
       y = "Frequency")
```

```{r}
# Log transformation (log(x) or log(x + 1) to handle zeros)
Band5_logvalues <- log(band5values + 1)  # Use log(x + 1) to avoid log(0)

ggplot(Band5_logvalues,
       aes(x= Band_1))+
  geom_histogram(bins = 100)+
  geom_vline(aes(xintercept = median(Band_1)), color = "red", linetype = "dashed", size = 0.75)+
  geom_vline(aes(xintercept = quantile(Band_1, 0.25)), color = "green", linetype = "dotted", size = 1) +
  geom_vline(aes(xintercept = quantile(Band_1, 0.75)), color = "blue", linetype = "dotted", size = 1) +
  labs(title = "5 Band", 
       x = "LOG Vulnerability", 
       y = "Frequency")
```

```{r}
# Log transformation (log(x) or log(x + 1) to handle zeros)
Band4_logvalues <- log(band4values + 1)  # Use log(x + 1) to avoid log(0)

ggplot(Band4_logvalues,
       aes(x= Band_1))+
  geom_histogram(bins = 100)+
  geom_vline(aes(xintercept = median(Band_1)), color = "red", linetype = "dashed", size = 0.75)+
  geom_vline(aes(xintercept = quantile(Band_1, 0.25)), color = "green", linetype = "dotted", size = 1) +
  geom_vline(aes(xintercept = quantile(Band_1, 0.75)), color = "blue", linetype = "dotted", size = 1) +
  labs(title = "4 Band", 
       x = "LOG Vulnerability", 
       y = "Frequency")
```

## For comparison, I am going to make panes which have the map, and both historgams 

first get all the log values made
```{r}
# Use log(x + 1) to avoid log(0)
Band1_logvalues <- log(band1values + 1)
Band2_logvalues <- log(band2values + 1)  
Band3_logvalues <- log(band3values + 1)  
Band4_logvalues <- log(band4values + 1)  
Band5_logvalues <- log(band5values + 1)
Band10_logvalues <- log(band10values + 1)
Band20_logvalues <- log(band20values + 1)
ExpBand4_logvalues <- log(expband4values + 1)
ExpBand5_logvalues <- log(expband5values + 1)
ExpBand6_logvalues <- log(expband6values + 1)
```

1band 
```{r}
#save the raster as a dataframe for ggplot to use
B1map_df <- as.data.frame(band1TIFF, xy = TRUE, na.rm = TRUE)
colnames(B1map_df)[3] <- "Vulnerability"
summary(B1map_df)

#ggplot it
B1map <- ggplot(B1map_df, aes(x = x, y = y, fill = Vulnerability)) +
  geom_raster() +  
  scale_fill_viridis_c(option = 'plasma', trans = 'log') + 
  coord_equal() +  # Ensures the aspect ratio is equal
  theme_minimal() +  
  labs(title = "1 Band Vulnerability Output", x = "Longitude", y = "Latitude")

B1Histogram<- ggplot(band1values,
       aes(x= Band_1))+
  geom_histogram(bins = 100)+
  geom_vline(aes(xintercept = median(Band_1)), color = "red", linetype = "dashed", size = 0.75)+
  geom_vline(aes(xintercept = quantile(Band_1, 0.25)), color = "green", linetype = "dotted", size = 1) +
  geom_vline(aes(xintercept = quantile(Band_1, 0.75)), color = "blue", linetype = "dotted", size = 1) +
  labs(title = "1 Band Vulnerability Estimate Distribution", 
       x = "Vulnerability", 
       y = "Frequency")

B1logHistorgam <- ggplot(Band1_logvalues,
       aes(x= Band_1))+
  geom_histogram(bins = 100)+
  geom_vline(aes(xintercept = median(Band_1)), color = "red", linetype = "dashed", size = 0.75)+
  geom_vline(aes(xintercept = quantile(Band_1, 0.25)), color = "green", linetype = "dotted", size = 1) +
  geom_vline(aes(xintercept = quantile(Band_1, 0.75)), color = "blue", linetype = "dotted", size = 1) +
  labs(title = "1 Band Log Vulnerability Estimate Distribution", 
       x = "LOG Vulnerability", 
       y = "Frequency")

# Open PNG device with ragg
agg_png("C:/Users/shil5756/Pictures/Saved Pictures/SensitivityAnalysisR/collatedband1.png", width = 6000, height = 1650, res = 300)

# Arrange the plots side by side
grid.arrange(B1map, B1Histogram, B1logHistorgam, ncol = 3)

# Close the PNG device to save the file
dev.off()
```

2band
```{r}
#save the raster as a dataframe for ggplot to use
B2map_df <- as.data.frame(band2TIFF, xy = TRUE, na.rm = TRUE)
colnames(B2map_df)[3] <- "Vulnerability"

#ggplot it
B2map <- ggplot(B2map_df, aes(x = x, y = y, fill = Vulnerability)) +
  geom_raster() +  
  scale_fill_viridis_c(option = 'plasma', trans = 'log') + 
  coord_equal() +  # Ensures the aspect ratio is equal
  theme_minimal() +  
  labs(title = "2 Band Vulnerability Output", x = "Longitude", y = "Latitude")

B2Histogram<- ggplot(band2values,
       aes(x= Band_1))+
  geom_histogram(bins = 100)+
  geom_vline(aes(xintercept = median(Band_1)), color = "red", linetype = "dashed", size = 0.75)+
  geom_vline(aes(xintercept = quantile(Band_1, 0.25)), color = "green", linetype = "dotted", size = 1) +
  geom_vline(aes(xintercept = quantile(Band_1, 0.75)), color = "blue", linetype = "dotted", size = 1) +
  labs(title = "2 Band Vulnerability Estimate Distribution", 
       x = "Vulnerability", 
       y = "Frequency")

B2logHistorgam <- ggplot(Band2_logvalues,
       aes(x= Band_1))+
  geom_histogram(bins = 100)+
  geom_vline(aes(xintercept = median(Band_1)), color = "red", linetype = "dashed", size = 0.75)+
  geom_vline(aes(xintercept = quantile(Band_1, 0.25)), color = "green", linetype = "dotted", size = 1) +
  geom_vline(aes(xintercept = quantile(Band_1, 0.75)), color = "blue", linetype = "dotted", size = 1) +
  labs(title = "2 Band Log Vulnerability Estimate Distribution", 
       x = "LOG Vulnerability", 
       y = "Frequency")

# Open PNG device with ragg
agg_png("C:/Users/shil5756/Pictures/Saved Pictures/SensitivityAnalysisR/collatedband2.png", width = 6000, height = 1650, res = 300)

# Arrange the plots side by side
grid.arrange(B2map, B2Histogram, B2logHistorgam, ncol = 3)

# Close the PNG device to save the file
dev.off()
```

3band
```{r}
#save the raster as a dataframe for ggplot to use
B3map_df <- as.data.frame(band3TIFF, xy = TRUE, na.rm = TRUE)
colnames(B3map_df)[3] <- "Vulnerability"

#ggplot it
B3map <- ggplot(B3map_df, aes(x = x, y = y, fill = Vulnerability)) +
  geom_raster() +  
  scale_fill_viridis_c(option = 'plasma', trans = 'log') + 
  coord_equal() +  # Ensures the aspect ratio is equal
  theme_minimal() +  
  labs(title = "3 Band Vulnerability Output", x = "Longitude", y = "Latitude")

B3Histogram<- ggplot(band3values,
       aes(x= Band_1))+
  geom_histogram(bins = 100)+
  geom_vline(aes(xintercept = median(Band_1)), color = "red", linetype = "dashed", size = 0.75)+
  geom_vline(aes(xintercept = quantile(Band_1, 0.25)), color = "green", linetype = "dotted", size = 1) +
  geom_vline(aes(xintercept = quantile(Band_1, 0.75)), color = "blue", linetype = "dotted", size = 1) +
  labs(title = "3 Band Vulnerability Estimate Distribution", 
       x = "Vulnerability", 
       y = "Frequency")

B3logHistorgam <- ggplot(Band3_logvalues,
       aes(x= Band_1))+
  geom_histogram(bins = 100)+
  geom_vline(aes(xintercept = median(Band_1)), color = "red", linetype = "dashed", size = 0.75)+
  geom_vline(aes(xintercept = quantile(Band_1, 0.25)), color = "green", linetype = "dotted", size = 1) +
  geom_vline(aes(xintercept = quantile(Band_1, 0.75)), color = "blue", linetype = "dotted", size = 1) +
  labs(title = "3 Band Log Vulnerability Estimate Distribution", 
       x = "LOG Vulnerability", 
       y = "Frequency")

# Open PNG device with ragg
agg_png("C:/Users/shil5756/Pictures/Saved Pictures/SensitivityAnalysisR/collatedband3.png", width = 6000, height = 1650, res = 300)

# Arrange the plots side by side
grid.arrange(B3map, B3Histogram, B3logHistorgam, ncol = 3)

# Close the PNG device to save the file
dev.off()
```

4band
```{r}
#save the raster as a dataframe for ggplot to use
B4map_df <- as.data.frame(band4TIFF, xy = TRUE, na.rm = TRUE)
colnames(B4map_df)[3] <- "Vulnerability"

#ggplot it
B4map <- ggplot(B4map_df, aes(x = x, y = y, fill = Vulnerability)) +
  geom_raster() +  
  scale_fill_viridis_c(option = 'plasma', trans = 'log') + 
  coord_equal() +  # Ensures the aspect ratio is equal
  theme_minimal() +  
  labs(title = "4 Band Vulnerability Output", x = "Longitude", y = "Latitude")

B4Histogram<- ggplot(band4values,
       aes(x= Band_1))+
  geom_histogram(bins = 100)+
  geom_vline(aes(xintercept = median(Band_1)), color = "red", linetype = "dashed", size = 0.75)+
  geom_vline(aes(xintercept = quantile(Band_1, 0.25)), color = "green", linetype = "dotted", size = 1) +
  geom_vline(aes(xintercept = quantile(Band_1, 0.75)), color = "blue", linetype = "dotted", size = 1) +
  labs(title = "4 Band Vulnerability Estimate Distribution", 
       x = "Vulnerability", 
       y = "Frequency")

B4logHistorgam <- ggplot(Band4_logvalues,
       aes(x= Band_1))+
  geom_histogram(bins = 100)+
  geom_vline(aes(xintercept = median(Band_1)), color = "red", linetype = "dashed", size = 0.75)+
  geom_vline(aes(xintercept = quantile(Band_1, 0.25)), color = "green", linetype = "dotted", size = 1) +
  geom_vline(aes(xintercept = quantile(Band_1, 0.75)), color = "blue", linetype = "dotted", size = 1) +
  labs(title = "4 Band Log Vulnerability Estimate Distribution", 
       x = "LOG Vulnerability", 
       y = "Frequency")

# Open PNG device with ragg
agg_png("C:/Users/shil5756/Pictures/Saved Pictures/SensitivityAnalysisR/collatedband4.png", width = 6000, height = 1650, res = 300)

# Arrange the plots side by side
grid.arrange(B4map, B4Histogram, B4logHistorgam, ncol = 3)

# Close the PNG device to save the file
dev.off()
```

5band
```{r}
#save the raster as a dataframe for ggplot to use
B5map_df <- as.data.frame(band5TIFF, xy = TRUE, na.rm = TRUE)
colnames(B5map_df)[3] <- "Vulnerability"

#ggplot it
B5map <- ggplot(B5map_df, aes(x = x, y = y, fill = Vulnerability)) +
  geom_raster() +  
  scale_fill_viridis_c(option = 'plasma', trans = 'log') + 
  coord_equal() +  # Ensures the aspect ratio is equal
  theme_minimal() +  
  labs(title = "5 Band Vulnerability Output", x = "Longitude", y = "Latitude")

B5Histogram<- ggplot(band5values,
       aes(x= Band_1))+
  geom_histogram(bins = 100)+
  geom_vline(aes(xintercept = median(Band_1)), color = "red", linetype = "dashed", size = 0.75)+
  geom_vline(aes(xintercept = quantile(Band_1, 0.25)), color = "green", linetype = "dotted", size = 1) +
  geom_vline(aes(xintercept = quantile(Band_1, 0.75)), color = "blue", linetype = "dotted", size = 1) +
  labs(title = "5 Band Vulnerability Estimate Distribution", 
       x = "Vulnerability", 
       y = "Frequency")

B5logHistorgam <- ggplot(Band5_logvalues,
       aes(x= Band_1))+
  geom_histogram(bins = 100)+
  geom_vline(aes(xintercept = median(Band_1)), color = "red", linetype = "dashed", size = 0.75)+
  geom_vline(aes(xintercept = quantile(Band_1, 0.25)), color = "green", linetype = "dotted", size = 1) +
  geom_vline(aes(xintercept = quantile(Band_1, 0.75)), color = "blue", linetype = "dotted", size = 1) +
  labs(title = "5 Band Log Vulnerability Estimate Distribution", 
       x = "LOG Vulnerability", 
       y = "Frequency")

# Open PNG device with ragg
agg_png("C:/Users/shil5756/Pictures/Saved Pictures/SensitivityAnalysisR/collatedband5.png", width = 6000, height = 1650, res = 300)

# Arrange the plots side by side
grid.arrange(B5map, B5Histogram, B5logHistorgam, ncol = 3)

# Close the PNG device to save the file
dev.off()
```

10band
```{r}
#save the raster as a dataframe for ggplot to use
B10map_df <- as.data.frame(band10TIFF, xy = TRUE, na.rm = TRUE)
colnames(B10map_df)[3] <- "Vulnerability"

#ggplot it
B10map <- ggplot(B10map_df, aes(x = x, y = y, fill = Vulnerability)) +
  geom_raster() +  
  scale_fill_viridis_c(option = 'plasma', trans = 'log') + 
  coord_equal() +  # Ensures the aspect ratio is equal
  theme_minimal() +  
  labs(title = "10 Band Vulnerability Output", x = "Longitude", y = "Latitude")

B10Histogram<- ggplot(band10values,
       aes(x= Band_1))+
  geom_histogram(bins = 100)+
  geom_vline(aes(xintercept = median(Band_1)), color = "red", linetype = "dashed", size = 0.75)+
  geom_vline(aes(xintercept = quantile(Band_1, 0.25)), color = "green", linetype = "dotted", size = 1) +
  geom_vline(aes(xintercept = quantile(Band_1, 0.75)), color = "blue", linetype = "dotted", size = 1) +
  labs(title = "10 Band Vulnerability Estimate Distribution", 
       x = "Vulnerability", 
       y = "Frequency")

B10logHistorgam <- ggplot(Band10_logvalues,
       aes(x= Band_1))+
  geom_histogram(bins = 100)+
  geom_vline(aes(xintercept = median(Band_1)), color = "red", linetype = "dashed", size = 0.75)+
  geom_vline(aes(xintercept = quantile(Band_1, 0.25)), color = "green", linetype = "dotted", size = 1) +
  geom_vline(aes(xintercept = quantile(Band_1, 0.75)), color = "blue", linetype = "dotted", size = 1) +
  labs(title = "10 Band Log Vulnerability Estimate Distribution", 
       x = "LOG Vulnerability", 
       y = "Frequency")

# Open PNG device with ragg
agg_png("C:/Users/shil5756/Pictures/Saved Pictures/SensitivityAnalysisR/collatedband10.png", width = 6000, height = 1650, res = 300)

# Arrange the plots side by side
grid.arrange(B10map, B10Histogram, B10logHistorgam, ncol = 3)

# Close the PNG device to save the file
dev.off()
```


20band
```{r}
#save the raster as a dataframe for ggplot to use
B20map_df <- as.data.frame(band20TIFF, xy = TRUE, na.rm = TRUE)
colnames(B20map_df)[3] <- "Vulnerability"

#ggplot it
B20map <- ggplot(B20map_df, aes(x = x, y = y, fill = Vulnerability)) +
  geom_raster() +  
  scale_fill_viridis_c(option = 'plasma', trans = 'log') + 
  coord_equal() +  # Ensures the aspect ratio is equal
  theme_minimal() +  
  labs(title = "20 Band Vulnerability Output", x = "Longitude", y = "Latitude")

B20Histogram<- ggplot(band20values,
       aes(x= Band_1))+
  geom_histogram(bins = 100)+
  geom_vline(aes(xintercept = median(Band_1)), color = "red", linetype = "dashed", size = 0.75)+
  geom_vline(aes(xintercept = quantile(Band_1, 0.25)), color = "green", linetype = "dotted", size = 1) +
  geom_vline(aes(xintercept = quantile(Band_1, 0.75)), color = "blue", linetype = "dotted", size = 1) +
  labs(title = "20 Band Vulnerability Estimate Distribution", 
       x = "Vulnerability", 
       y = "Frequency")

B20logHistorgam <- ggplot(Band20_logvalues,
       aes(x= Band_1))+
  geom_histogram(bins = 100)+
  geom_vline(aes(xintercept = median(Band_1)), color = "red", linetype = "dashed", size = 0.75)+
  geom_vline(aes(xintercept = quantile(Band_1, 0.25)), color = "green", linetype = "dotted", size = 1) +
  geom_vline(aes(xintercept = quantile(Band_1, 0.75)), color = "blue", linetype = "dotted", size = 1) +
  labs(title = "20 Band Log Vulnerability Estimate Distribution", 
       x = "LOG Vulnerability", 
       y = "Frequency")

# Open PNG device with ragg
agg_png("C:/Users/shil5756/Pictures/Saved Pictures/SensitivityAnalysisR/collatedband20.png", width = 6000, height = 1650, res = 300)

# Arrange the plots side by side
grid.arrange(B20map, B20Histogram, B20logHistorgam, ncol = 3)

# Close the PNG device to save the file
dev.off()
```

exp4band
```{r}
#save the raster as a dataframe for ggplot to use
expB4map_df <- as.data.frame(expband4TIFF, xy = TRUE, na.rm = TRUE)
colnames(expB4map_df)[3] <- "Vulnerability"

#ggplot it
expB4map <- ggplot(expB4map_df, aes(x = x, y = y, fill = Vulnerability)) +
  geom_raster() +  
  scale_fill_viridis_c(option = 'plasma', trans = 'log') + 
  coord_equal() +  # Ensures the aspect ratio is equal
  theme_minimal() +  
  labs(title = "Exponential 4 Band Vulnerability Output", x = "Longitude", y = "Latitude")

expB4Histogram<- ggplot(expband4values,
       aes(x= Band_1))+
  geom_histogram(bins = 100)+
  geom_vline(aes(xintercept = median(Band_1)), color = "red", linetype = "dashed", size = 0.75)+
  geom_vline(aes(xintercept = quantile(Band_1, 0.25)), color = "green", linetype = "dotted", size = 1) +
  geom_vline(aes(xintercept = quantile(Band_1, 0.75)), color = "blue", linetype = "dotted", size = 1) +
  labs(title = "Exponential 4 Band Vulnerability Estimate Distribution", 
       x = "Vulnerability", 
       y = "Frequency")

expB4logHistorgam <- ggplot(ExpBand4_logvalues,
       aes(x= Band_1))+
  geom_histogram(bins = 100)+
  geom_vline(aes(xintercept = median(Band_1)), color = "red", linetype = "dashed", size = 0.75)+
  geom_vline(aes(xintercept = quantile(Band_1, 0.25)), color = "green", linetype = "dotted", size = 1) +
  geom_vline(aes(xintercept = quantile(Band_1, 0.75)), color = "blue", linetype = "dotted", size = 1) +
  labs(title = "Exponential 4 Band Log Vulnerability Estimate Distribution", 
       x = "LOG Vulnerability", 
       y = "Frequency")

# Open PNG device with ragg
agg_png("C:/Users/shil5756/Pictures/Saved Pictures/SensitivityAnalysisR/collatedEXPband4.png", width = 6000, height = 1650, res = 300)

# Arrange the plots side by side
grid.arrange(expB4map, expB4Histogram, expB4logHistorgam, ncol = 3)

# Close the PNG device to save the file
dev.off()
```


exp5band
```{r}
#save the raster as a dataframe for ggplot to use
expB5map_df <- as.data.frame(expband5TIFF, xy = TRUE, na.rm = TRUE)
colnames(expB5map_df)[3] <- "Vulnerability"

#ggplot it
expB5map <- ggplot(expB5map_df, aes(x = x, y = y, fill = Vulnerability)) +
  geom_raster() +  
  scale_fill_viridis_c(option = 'plasma', trans = 'log') + 
  coord_equal() +  # Ensures the aspect ratio is equal
  theme_minimal() +  
  labs(title = "Exponential 5 Band Vulnerability Output", x = "Longitude", y = "Latitude")

expB5Histogram<- ggplot(expband5values,
       aes(x= Band_1))+
  geom_histogram(bins = 100)+
  geom_vline(aes(xintercept = median(Band_1)), color = "red", linetype = "dashed", size = 0.75)+
  geom_vline(aes(xintercept = quantile(Band_1, 0.25)), color = "green", linetype = "dotted", size = 1) +
  geom_vline(aes(xintercept = quantile(Band_1, 0.75)), color = "blue", linetype = "dotted", size = 1) +
  labs(title = "Exponential 5 Band Vulnerability Estimate Distribution", 
       x = "Vulnerability", 
       y = "Frequency")

expB5logHistorgam <- ggplot(ExpBand5_logvalues,
       aes(x= Band_1))+
  geom_histogram(bins = 100)+
  geom_vline(aes(xintercept = median(Band_1)), color = "red", linetype = "dashed", size = 0.75)+
  geom_vline(aes(xintercept = quantile(Band_1, 0.25)), color = "green", linetype = "dotted", size = 1) +
  geom_vline(aes(xintercept = quantile(Band_1, 0.75)), color = "blue", linetype = "dotted", size = 1) +
  labs(title = "Exponential 5 Band Log Vulnerability Estimate Distribution", 
       x = "LOG Vulnerability", 
       y = "Frequency")

# Open PNG device with ragg
agg_png("C:/Users/shil5756/Pictures/Saved Pictures/SensitivityAnalysisR/collatedEXPband5.png", width = 6000, height = 1650, res = 300)

# Arrange the plots side by side
grid.arrange(expB5map, expB5Histogram, expB5logHistorgam, ncol = 3)

# Close the PNG device to save the file
dev.off()
```

exp6band
```{r}
#save the raster as a dataframe for ggplot to use
expB6map_df <- as.data.frame(expband6TIFF, xy = TRUE, na.rm = TRUE)
colnames(expB6map_df)[3] <- "Vulnerability"

#ggplot it
expB6map <- ggplot(expB6map_df, aes(x = x, y = y, fill = Vulnerability)) +
  geom_raster() +  
  scale_fill_viridis_c(option = 'plasma', trans = 'log') + 
  coord_equal() +  # Ensures the aspect ratio is equal
  theme_minimal() +  
  labs(title = "Exponential 6 Band Vulnerability Output", x = "Longitude", y = "Latitude")

expB6Histogram<- ggplot(expband6values,
       aes(x= Band_1))+
  geom_histogram(bins = 100)+
  geom_vline(aes(xintercept = median(Band_1)), color = "red", linetype = "dashed", size = 0.75)+
  geom_vline(aes(xintercept = quantile(Band_1, 0.25)), color = "green", linetype = "dotted", size = 1) +
  geom_vline(aes(xintercept = quantile(Band_1, 0.75)), color = "blue", linetype = "dotted", size = 1) +
  labs(title = "Exponential 6 Band Vulnerability Estimate Distribution", 
       x = "Vulnerability", 
       y = "Frequency")

expB6logHistorgam <- ggplot(ExpBand6_logvalues,
       aes(x= Band_1))+
  geom_histogram(bins = 100)+
  geom_vline(aes(xintercept = median(Band_1)), color = "red", linetype = "dashed", size = 0.75)+
  geom_vline(aes(xintercept = quantile(Band_1, 0.25)), color = "green", linetype = "dotted", size = 1) +
  geom_vline(aes(xintercept = quantile(Band_1, 0.75)), color = "blue", linetype = "dotted", size = 1) +
  labs(title = "Exponential 6 Band Log Vulnerability Estimate Distribution", 
       x = "LOG Vulnerability", 
       y = "Frequency")

# Open PNG device with ragg
agg_png("C:/Users/shil5756/Pictures/Saved Pictures/SensitivityAnalysisR/collatedEXPband6.png", width = 6000, height = 1650, res = 300)

# Arrange the plots side by side
grid.arrange(expB6map, expB6Histogram, expB6logHistorgam, ncol = 3)

# Close the PNG device to save the file
dev.off()
```















