---
title: "SensitivityExploration"
author: "Kristian Suszczenia"
date: "2025-01-20"
output: html_document
---

This code entails the sensitivty analysis and graphs produced in the supplement of our study. This analysis aims to produce histograms of the data produced by the different arbitrary strength banding, as well as plot computaiton time. As with all the repository, #explanations will only be shown on first instance of the code's use.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Load required packages
```{r load pakcages}
#use install.packages() if needed
library(terra)
library(ggplot2)
library(gridExtra)
library(ragg)
library(here)
```

#functions

```{r Load functions}
#this function was used to save png images, though saving is not shown in this public facing code
save_figure <- function(Figure_to_save,
                                  filename, width, height, res, scaling){
  agg_png(filename, 
          width   =  width, 
          height  =  height, 
          units   =  "cm", 
          res     =  res, 
          scaling =  scaling)
  Figure_to_save
  print(Figure_to_save)
  dev.off()
}
```

### Part 1: create the data

##Import the band 1 metric to test
I am importing as tif

```{r Load Band 1}
#load the raster of distrotion vulnerability for each banding structure on the same species (LWshort)
band1TIFF <- rast(here('DistortionMetricDesign', 'SensitivityAnalysis', 'BandingTest_Outputs', 'Band1.tif'))

#you can look at the object
band1TIFF
#get sumary stats
summary(band1TIFF)
#and plot it easily!
plot(band1TIFF)
```

But, mainly, we want to save the mean, median, and 2SD
```{r Calculate Stats, Band1}
#clean NAs
band1values <- na.omit(values(band1TIFF))
#calculate and save mean
band1mean <- mean(band1values) # = 1.463686
#calculate and save median
band1median <- median(band1values) # = 0
#calculate and save SD
band1SD <- sd(band1values) # = 6.831418

#and add the computation time, sourced from ArcGis tool report
band1time <- 40
```

#Repeat to get values for all the bands

2 band
```{r LOad Band 2}
band2TIFF <- rast(here('DistortionMetricDesign', 'SensitivityAnalysis', 'BandingTest_Outputs', 'band2.tif'))
band2values <- na.omit(values(band2TIFF))
band2mean <- mean(band2values) # = 1.970988
band2median <- median(band2values) # = 0
band2SD <- sd(band2values) # = 7.061939
band2time <- 70
```

3 band
```{r Load Band 3}
band3TIFF <- rast(here('DistortionMetricDesign', 'SensitivityAnalysis', 'BandingTest_Outputs', 'band3.tif'))
band3values <- na.omit(values(band3TIFF))
band3mean <- mean(band3values) # = 4.8856
band3median <- median(band3values) # = 0.7585
band3SD <- sd(band3values) # = 12.9285
band3time <- 77
```

4band
```{r Load band 4}
band4TIFF <- rast(here('DistortionMetricDesign', 'SensitivityAnalysis', 'BandingTest_Outputs', 'band4.tif'))
band4values <- na.omit(values(band4TIFF))
band4mean <- mean(band4values) # = 6.3546
band4median <- median(band4values) # = 1.4470
band4SD <- sd(band4values) # = 14.1165
band4time <- 97
```

5band
```{r load band 5}
band5TIFF <- rast(here('DistortionMetricDesign', 'SensitivityAnalysis', 'BandingTest_Outputs', 'band5.tif'))
band5values <- na.omit(values(band5TIFF))
band5mean <- mean(band5values) 
band5median <- median(band5values) 
band5SD <- sd(band5values)
band5time <- 119
```

10band
```{r load band 10}
band10TIFF <- rast(here('DistortionMetricDesign', 'SensitivityAnalysis', 'BandingTest_Outputs', 'band10.tif'))
band10values <- na.omit(values(band10TIFF))
band10mean <- mean(band10values) 
band10median <- median(band10values) 
band10SD <- sd(band10values) 
band10time <- 177
```

20band
```{r load band 20}
band20TIFF <- rast(here('DistortionMetricDesign', 'SensitivityAnalysis', 'BandingTest_Outputs', '20band.tif'))
band20values <- na.omit(values(band20TIFF))
band20mean <- mean(band20values) 
band20median <- median(band20values) 
band20SD <- sd(band20values) 
band20time <- 370
```

exponential 4 band
```{r load 4 band exponential}
expband4TIFF <- rast(here('DistortionMetricDesign', 'SensitivityAnalysis', 'BandingTest_Outputs', 'exp4band.tif'))
expband4values <- na.omit(values(expband4TIFF))
expband4mean <- mean(expband4values) 
expband4median <- median(expband4values) 
expband4SD <- sd(expband4values) 
expband4time <- 99
```

exponential 5 band
```{r load 5 band exponential}
expband5TIFF <- rast(here('DistortionMetricDesign', 'SensitivityAnalysis', 'BandingTest_Outputs', 'exp5band.tif'))
expband5values <- na.omit(values(expband5TIFF))
expband5mean <- mean(expband5values) 
expband5median <- median(expband5values) 
expband5SD <- sd(expband5values) 
expband5time <- 138
```

exponential 6 band
```{r load 6 band exponential}
expband6TIFF <- rast(here('DistortionMetricDesign', 'SensitivityAnalysis', 'BandingTest_Outputs', 'exp6band.tif'))
expband6values <- na.omit(values(expband6TIFF))
expband6mean <- mean(expband6values) 
expband6median <- median(expband6values) 
expband6SD <- sd(expband6values) 
expband6time <- 148
```

#collate the data into a single object so we can plot it
```{r Create Single Dataframe}
#Build each column using c()
names <- c('1band', '2band', '3band', '4band', '5band', '10band', '20band', 'exp4band', 'exp5band', 'exp6band')
means <- c(band1mean, band2mean, band3mean, band4mean, band5mean, band10mean, band20mean, expband4mean, expband5mean, expband6mean)
medians <- c(band1median, band2median, band3median, band4median, band5median, band10median, band20median, expband4median, expband5median, expband6median)
bandnum <- c(1, 2, 3, 4, 5, 10, 20, 4, 5, 6)
comptime <- c(band1time, band2time, band3time, band4time, band5time, band10time, band20time, expband4time, expband5time, expband6time)
metrictype <- c('linear', 'linear', 'linear', 'linear', 'linear', 'linear', 'linear', 'exponential', 'exponential', 'exponential')

#Build the dataframe by combining columns
BandStatsDf <- data.frame(Bandname= names, mean= means, median= medians, bands= bandnum, computation_time= comptime, metric_type= metrictype)

#Set the catagorical variables as catagorical (as 'factor')
BandStatsDf$metric_type <- factor(BandStatsDf$metric_type, levels = c('exponential', 'linear'))
```

### Part 2: Comparison Plots

#plot the means against number of bands

```{r Plot means v banding}
#plot means v banding
ggplot(BandStatsDf, 
       aes(x = bands, 
           y = mean, 
           colour = metric_type)) + #colour by the type; exponential or linear
  geom_point(size = 3) + 
  labs(title = "Sensitivty Analysis of Number of strength bands", 
       x = "Number of Bands", 
       y = "Mean Vulnerability Estimate") +
  theme_linedraw()

#save it if desired
```

# plot the medians against number of bands

```{r plot medians v banding}
ggplot(BandStatsDf, 
       aes(x = bands, 
           y = median, 
           colour = metric_type)) +
  geom_point(size = 3) + 
  labs(title = "Sensitivty Analysis of Number of strength bands", 
       x = "Number of Bands", 
       y = "Median Vulnerability Estimate") +
  theme_minimal()
```

#plot the computation time against number of bands

```{r plot computation v banding}
ggplot(BandStatsDf, 
       aes(x = bands, 
           y = computation_time, 
           colour = metric_type)) +
  geom_point(size = 3) + 
  labs(title = "Sensitivty Analysis of Number of strength bands", 
       x = "Number of Bands", 
       y = "Computation Time") +
  theme_minimal()
```

#plot the computation time against means

```{r plot computation v means}
ggplot(BandStatsDf, 
       aes(x = mean, 
           y = computation_time, 
           colour = metric_type,
           size = bands)) + #here we also want to display band number somehow, here I use size
  geom_point() + 
  labs(title = "Sensitivty Analysis of Number of strength bands", 
       x = "Mean Vulnerability Estimate", 
       y = "Computation Time") +
  theme_minimal()
```

### PART 3: Histograms to understand the data

#Explore Histogram

1 Band
```{r Hist, band 1}
#view the data
summary(band1values)

#plot historgram
ggplot(band1values,
       aes(x= Band_1))+
  geom_histogram(bins = 100)+ #set the number of bins to aid viewing, but be consistant
  geom_vline(aes(xintercept = median(Band_1)), color = "red", linetype = "dashed", size = 0.75)+ #include the median and interquartile range
  geom_vline(aes(xintercept = quantile(Band_1, 0.25)), color = "green", linetype = "dotted", size = 1) +
  geom_vline(aes(xintercept = quantile(Band_1, 0.75)), color = "blue", linetype = "dotted", size = 1) +
  labs(title = "1 Band", 
       x = "Vulnerability", 
       y = "Frequency")
```

2 Band
```{r Hist, band 2}
summary(band2values)

ggplot(band2values,
       aes(x= Band_1))+
  geom_histogram(bins = 100)+
  geom_vline(aes(xintercept = median(Band_1)), color = "red", linetype = "dashed", size = 0.75)+
  geom_vline(aes(xintercept = quantile(Band_1, 0.25)), color = "green", linetype = "dotted", size = 1) +
  geom_vline(aes(xintercept = quantile(Band_1, 0.75)), color = "blue", linetype = "dotted", size = 1) +
  labs(title = "2 Band", 
       x = "Vulnerability", 
       y = "Frequency")
```

3 Band
```{r Hist, band 3}
summary(band3values)

ggplot(band3values,
       aes(x= Band_1))+
  geom_histogram(bins = 100)+
  geom_vline(aes(xintercept = median(Band_1)), color = "red", linetype = "dashed", size = 0.75)+
  geom_vline(aes(xintercept = quantile(Band_1, 0.25)), color = "green", linetype = "dotted", size = 1) +
  geom_vline(aes(xintercept = quantile(Band_1, 0.75)), color = "blue", linetype = "dotted", size = 1) +
  labs(title = "3 Band", 
       x = "Vulnerability", 
       y = "Frequency")
```

4 Band
```{r Hist, band 4}
summary(band4values)

ggplot(band4values,
       aes(x= Band_1))+
  geom_histogram(bins = 100)+
  geom_vline(aes(xintercept = median(Band_1)), color = "red", linetype = "dashed", size = 0.75)+
  geom_vline(aes(xintercept = quantile(Band_1, 0.25)), color = "green", linetype = "dotted", size = 1) +
  geom_vline(aes(xintercept = quantile(Band_1, 0.75)), color = "blue", linetype = "dotted", size = 1) +
  labs(title = "4 Band", 
       x = "Vulnerability", 
       y = "Frequency")
```

5 Band
```{r Hist, band 5}
summary(band5values)

ggplot(band5values,
       aes(x= Band_1))+
  geom_histogram(bins = 100)+
  geom_vline(aes(xintercept = median(Band_1)), color = "red", linetype = "dashed", size = 0.75)+
  geom_vline(aes(xintercept = quantile(Band_1, 0.25)), color = "green", linetype = "dotted", size = 1) +
  geom_vline(aes(xintercept = quantile(Band_1, 0.75)), color = "blue", linetype = "dotted", size = 1) +
  labs(title = "5 Band", 
       x = "Vulnerability", 
       y = "Frequency")
```

10 Band
```{r Hist, band 10}
summary(band10values)

ggplot(band10values,
       aes(x= Band_1))+
  geom_histogram(bins = 100)+
  geom_vline(aes(xintercept = median(Band_1)), color = "red", linetype = "dashed", size = 0.75)+
  geom_vline(aes(xintercept = quantile(Band_1, 0.25)), color = "green", linetype = "dotted", size = 1) +
  geom_vline(aes(xintercept = quantile(Band_1, 0.75)), color = "blue", linetype = "dotted", size = 1) +
  labs(title = "10 Band", 
       x = "Vulnerability", 
       y = "Frequency")
```

20 Band
```{r Hist, band 20}
summary(band20values)

ggplot(band20values,
       aes(x= Band_1))+
  geom_histogram(bins = 100)+
  geom_vline(aes(xintercept = median(Band_1)), color = "red", linetype = "dashed", size = 0.75)+
  geom_vline(aes(xintercept = quantile(Band_1, 0.25)), color = "green", linetype = "dotted", size = 1) +
  geom_vline(aes(xintercept = quantile(Band_1, 0.75)), color = "blue", linetype = "dotted", size = 1) +
  labs(title = "20 Band", 
       x = "Vulnerability", 
       y = "Frequency")
```

Exponential 4 Band
```{r Hist, band 4 exponential}
summary(expband4values)

ggplot(expband4values,
       aes(x= Band_1))+
  geom_histogram(bins = 100)+
  geom_vline(aes(xintercept = median(Band_1)), color = "red", linetype = "dashed", size = 0.75)+
  geom_vline(aes(xintercept = quantile(Band_1, 0.25)), color = "green", linetype = "dotted", size = 1) +
  geom_vline(aes(xintercept = quantile(Band_1, 0.75)), color = "blue", linetype = "dotted", size = 1) +
  labs(title = "Exponential 4 Band", 
       x = "Vulnerability", 
       y = "Frequency")
```

Exponential 5 Band
```{r Hist, band 5 exponential}
summary(expband5values)

ggplot(expband5values,
       aes(x= Band_1))+
  geom_histogram(bins = 100)+
  geom_vline(aes(xintercept = median(Band_1)), color = "red", linetype = "dashed", size = 0.75)+
  geom_vline(aes(xintercept = quantile(Band_1, 0.25)), color = "green", linetype = "dotted", size = 1) +
  geom_vline(aes(xintercept = quantile(Band_1, 0.75)), color = "blue", linetype = "dotted", size = 1) +
  labs(title = "Exponential 5 Band", 
       x = "Vulnerability", 
       y = "Frequency")
```

Exponential 6 Band
```{r Hist, band 6 exponential}
summary(expband6values)

ggplot(expband6values,
       aes(x= Band_1))+
  geom_histogram(bins = 100)+
  geom_vline(aes(xintercept = median(Band_1)), color = "red", linetype = "dashed", size = 0.75)+
  geom_vline(aes(xintercept = quantile(Band_1, 0.25)), color = "green", linetype = "dotted", size = 1) +
  geom_vline(aes(xintercept = quantile(Band_1, 0.75)), color = "blue", linetype = "dotted", size = 1) +
  labs(title = "Exponential 6 Band", 
       x = "Vulnerability", 
       y = "Frequency")
```

## Experiment with transformations

there is clearly some influential thing happening with banding structure, but it is hard to see. so we will plot the log response histogram too. 

```{r Log Hist, band 5 exponential}
# Log transformation (log(x) or log(x + 1) to handle zeros)
ExpBand5_logvalues <- log(expband5values + 1)  # Use log(x + 1) to avoid log(0)

#plot as before
ggplot(ExpBand5_logvalues,
       aes(x= Band_1))+
  geom_histogram(bins = 100)+
  geom_vline(aes(xintercept = median(Band_1)), color = "red", linetype = "dashed", size = 0.75)+
  geom_vline(aes(xintercept = quantile(Band_1, 0.25)), color = "green", linetype = "dotted", size = 1) +
  geom_vline(aes(xintercept = quantile(Band_1, 0.75)), color = "blue", linetype = "dotted", size = 1) +
  labs(title = "Exponential 6 Band", 
       x = "LOG Vulnerability", 
       y = "Frequency")
```

```{r Log Hist, band 6 exponential}
# Log transformation (log(x) or log(x + 1) to handle zeros)
ExpBand6_logvalues <- log(expband6values + 1)  # Use log(x + 1) to avoid log(0)

ggplot(ExpBand6_logvalues,
       aes(x= Band_1))+
  geom_histogram(bins = 100)+
  geom_vline(aes(xintercept = median(Band_1)), color = "red", linetype = "dashed", size = 0.75)+
  geom_vline(aes(xintercept = quantile(Band_1, 0.25)), color = "green", linetype = "dotted", size = 1) +
  geom_vline(aes(xintercept = quantile(Band_1, 0.75)), color = "blue", linetype = "dotted", size = 1) +
  labs(title = "Exponential 6 Band", 
       x = "LOG Vulnerability", 
       y = "Frequency")
```

```{r Log Hist, band 10}
# Log transformation (log(x) or log(x + 1) to handle zeros)
Band10_logvalues <- log(band10values + 1)  # Use log(x + 1) to avoid log(0)

ggplot(Band10_logvalues,
       aes(x= Band_1))+
  geom_histogram(bins = 100)+
  geom_vline(aes(xintercept = median(Band_1)), color = "red", linetype = "dashed", size = 0.75)+
  geom_vline(aes(xintercept = quantile(Band_1, 0.25)), color = "green", linetype = "dotted", size = 1) +
  geom_vline(aes(xintercept = quantile(Band_1, 0.75)), color = "blue", linetype = "dotted", size = 1) +
  labs(title = "20 Band", 
       x = "LOG Vulnerability", 
       y = "Frequency")
```

```{r Log Hist, band 20}
# Log transformation (log(x) or log(x + 1) to handle zeros)
Band20_logvalues <- log(band20values + 1)  # Use log(x + 1) to avoid log(0)

ggplot(Band20_logvalues,
       aes(x= Band_1))+
  geom_histogram(bins = 100)+
  geom_vline(aes(xintercept = median(Band_1)), color = "red", linetype = "dashed", size = 0.75)+
  geom_vline(aes(xintercept = quantile(Band_1, 0.25)), color = "green", linetype = "dotted", size = 1) +
  geom_vline(aes(xintercept = quantile(Band_1, 0.75)), color = "blue", linetype = "dotted", size = 1) +
  labs(title = "20 Band", 
       x = "LOG Vulnerability", 
       y = "Frequency")
```

```{r Log Hist, band 5}
# Log transformation (log(x) or log(x + 1) to handle zeros)
Band5_logvalues <- log(band5values + 1)  # Use log(x + 1) to avoid log(0)

ggplot(Band5_logvalues,
       aes(x= Band_1))+
  geom_histogram(bins = 100)+
  geom_vline(aes(xintercept = median(Band_1)), color = "red", linetype = "dashed", size = 0.75)+
  geom_vline(aes(xintercept = quantile(Band_1, 0.25)), color = "green", linetype = "dotted", size = 1) +
  geom_vline(aes(xintercept = quantile(Band_1, 0.75)), color = "blue", linetype = "dotted", size = 1) +
  labs(title = "5 Band", 
       x = "LOG Vulnerability", 
       y = "Frequency")
```

```{r Log Hist, band 4}
# Log transformation (log(x) or log(x + 1) to handle zeros)
Band4_logvalues <- log(band4values + 1)  # Use log(x + 1) to avoid log(0)

ggplot(Band4_logvalues,
       aes(x= Band_1))+
  geom_histogram(bins = 100)+
  geom_vline(aes(xintercept = median(Band_1)), color = "red", linetype = "dashed", size = 0.75)+
  geom_vline(aes(xintercept = quantile(Band_1, 0.25)), color = "green", linetype = "dotted", size = 1) +
  geom_vline(aes(xintercept = quantile(Band_1, 0.75)), color = "blue", linetype = "dotted", size = 1) +
  labs(title = "4 Band", 
       x = "LOG Vulnerability", 
       y = "Frequency")
```

For conclusions from this analysis, see the supplement of the dissertation paper. 








